---
title: "MitePopGen"
author: "Errbii M."
date: "2025-02-07"
output: html_document
---

please not that I have uploaded all the datasets to the data folder. To make plots and so on, you might need to adjust the path to the location of the data file in question.

```{r}
library(ggpubr)
library(ggtree)
library(patchwork)
library(gtools)
library(tidyverse)
library(data.table)
library(reshape2)
library(ggrepel)
library(ggbeeswarm)
library(ape)
library(caper)
```

## SNP filtering

Use VCFtools output files to visualize and decide for thresholds for various SNP metrics.

#### *allele frequency*
```{r}
# Visualize Allele frquencies, coverage and missingness:

# list of species to loop over:
species <- c("Ass", "Hga", "Hls", "Hrs", "Hti", "Ppr", "Sms") 

# load Freq
listOfAllFreqs<-list()


for (sp in species) { 
  file_path <- paste0("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/QCs/",sp,".varSNPs.biallpass.hom_QC/",sp,".varSNPs.biallpass.hom.freq.frq")
  
  if (file.exists(file_path)) {  # Check if file exists before reading
    listOfAllFreqs[[sp]] <- read.csv(file_path, sep='\t', header=TRUE)  # Read file
    listOfAllFreqs[[sp]]$species <- sp
    
  } else {
    warning(paste("File not found:", file_path))  # Print a warning if file is missing
  }
}

# Combine all tables into a single data frame by row binding
final_combined_freq <- bind_rows(listOfAllFreqs)

# set colors
cols <-c(
  "Ref_Allele"="#ff6666",
  "Alt_Allele"="#00cc44")

# Get unique species
species_list <- unique(final_combined_freq$species)

# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(sp) {
  species_data <- final_combined_freq %>% filter(species == sp)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = Ref_Allele, after_stat(density), fill = "Ref_Allele"), binwidth = 0.05, alpha = 0.5)+
    geom_histogram(aes(x = Alt_Allele, after_stat(density), fill = "Alt_Allele"), binwidth = 0.05, alpha =0.5)+
    labs(title = "Allele frequency", x = "Frequency", y = "Density")+
    scale_fill_manual(name="Allele",values=cols)+ 
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(sp)# Title for each species
  
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "bottom")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/QCs/Freq.pdf",width=9,height=6)


# clean working space:
rm(list = names(which(!unlist(eapply(.GlobalEnv, \(x) inherits(x, what = "function"))))))

```

#### *mean depth per site*

```{r depth,  echo=FALSE, message=FALSE, warning=FALSE}

# list of species to loop over:
species <- c("Ass", "Hga", "Hls", "Hrs", "Hti", "Ppr", "Sms")

# load Freq
listOfAllmeandepth<-list()


for (sp in species) { 
  file_path <- paste0("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/QCs/",sp,".varSNPs.biallpass.hom_QC/",sp,".varSNPs.biallpass.hom.site.depth.ldepth.mean")
  
  if (file.exists(file_path)) {  # Check if file exists before reading
    listOfAllmeandepth[[sp]] <- read.csv(file_path, sep='\t', header=TRUE)  # Read file
    listOfAllmeandepth[[sp]]$species <- sp
    
  } else {
    warning(paste("File not found:", file_path))  # Print a warning if file is missing
  }
}

# Combine all tables into a single data frame by row binding
final_combined_meandepth <- bind_rows(listOfAllmeandepth)

# Get unique species
species_list <- unique(final_combined_meandepth$species)

# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(sp) {
  species_data <- final_combined_meandepth %>% filter(species == sp)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = MEAN_DEPTH, stat(density)), binwidth = 1, fill = "grey")+ 
    labs(title = "Mean depth per site")+
    xlim(c(0,200))+
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(sp, ": ", quantile(species_data$MEAN_DEPTH, 0.05, na.rm = T), " - ", quantile(species_data$MEAN_DEPTH, 0.95, na.rm = T))) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "bottom")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/QCs/Depth.pdf",width=9,height=6)

# clean working space:
rm(list = names(which(!unlist(eapply(.GlobalEnv, \(x) inherits(x, what = "function"))))))

```
#### *missing data: SNP*

```{r missingness,  echo=FALSE, message=FALSE, warning=FALSE}

# list of species to loop over:
species <- c("Ass", "Hga", "Hls", "Hrs", "Hti", "Ppr", "Sms")

# load Freq
listOfAllmiss<-list()


for (sp in species) { 
  file_path <- paste0("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/QCs/",sp,".varSNPs.biallpass.hom_QC/",sp,".varSNPs.biallpass.hom.missing.site.lmiss")
  
  if (file.exists(file_path)) {  # Check if file exists before reading
    listOfAllmiss[[sp]] <- read.csv(file_path, sep='\t', header=TRUE)  # Read file
    listOfAllmiss[[sp]]$species <- sp
    
  } else {
    warning(paste("File not found:", file_path))  # Print a warning if file is missing
  }
}

# Combine all tables into a single data frame by row binding
final_combined_miss <- bind_rows(listOfAllmiss)

# Get unique species
species_list <- unique(final_combined_miss$species)

# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(sp) {
  species_data <- final_combined_miss %>% filter(species == sp)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = F_MISS, stat(density)),  binwidth = 0.1, fill = "grey")+
    labs(title = "Missingness on a per-site basis", x = "Frequency (binwidth = 0.1)", y = "Density")+
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(sp)) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "bottom")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/QCs/Miss.pdf",width=9,height=6)

# clean working space:
rm(list = names(which(!unlist(eapply(.GlobalEnv, \(x) inherits(x, what = "function"))))))

```

#### *missing data: INDV*

Just do `for i in */*SNPspass.biall.recode.missing.indv.imiss; do echo ${i}; cat ${i};done` in bash and check if there are any outliers.


#### *ME analysis: SFS plots*

```{r SFSs, echo=FALSE, message=FALSE, warning=FALSE}

# load all needed libraries:
library(tidyverse)
library(jcolors)
library(grid)

# devtools::install_github("jaredhuling/jcolors")

# Load genotype data:
gt <- read_delim("~/sciebo/Postdoc/sexAsex/popGen/sfs/Hrs.SFS.N7.txt", delim = "\t", col_names = F)
colnames(gt) <- c("occ", "profile", "alt")
head(gt)


# annotate data: add order based on frequency of occurrence:
df <- gt %>% group_by(alt) %>% mutate(ID = as.numeric(row_number()))


# show only N (here 20) most frequent genotypes:
df <- mutate(as_tibble(df), ID = ifelse(ID < 21, ID, "others"), profile = ifelse(ID == "others", "others", profile))

# group and count other:
df2 <- df %>% group_by(alt, profile, ID) %>% summarise(occ = sum(occ))

df2$ID <- ifelse((nchar(df2$ID) ==1) ,paste0("0",df2$ID),df2$ID)

sum(df2$occ)

# add NAs if count is less than N most frequent genotypes (here 20):
# Group by 'alt' and count the number of rows
result <- df2 %>%
  group_by(alt) %>%
  summarise(n_rows = n()) %>%
  ungroup() %>%
  mutate(n_add_rows = ifelse(n_rows < 21, 21 - n_rows, 0)) %>%
  mutate(add_rows = ifelse(n_rows < 21, TRUE, FALSE))

# Print the result
result[result$add_rows=="TRUE",]

# add rows
# eur
df2 <- df2 %>% ungroup() %>% 
  add_row(alt = c(replicate(14, 1), replicate(14, 13)), # n_add_rows, alt
          profile = c(replicate(13, "N/A"),"others", replicate(13, "N/A"),"others"), # n_add_rows - 1
          ID = c(seq(8,20,1),"others", seq(8,20,1),"others"), # n_rows + 1
          occ = c(replicate(14, 0), replicate(14, 0))) # same as alt


# order data:
df2$ID <- ifelse((nchar(df2$ID) ==1) ,paste0("0",df2$ID),df2$ID)

final <- df2[,c(1,4,3,2)] %>% 
  as.data.frame() %>% 
  arrange(alt, ID) 

# save this file:
write_delim(final, "~/sciebo/Postdoc/sexAsex/popGen/sfs/Hrs.SFS.N7.forR.ed.txt", delim = "\t")

# THEN GO AN USE THE SCRIPT INSIDE SRC OR RUN THE FOLLOWING SECTION

```

```{r SFSs, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(grid)
library(jcolors)

# Define file path
sfs_dir <- "~/sciebo/Postdoc/sexAsex/popGen/sfs/"

# List all files matching pattern *ed.txt
sfs_files <- list.files(path = sfs_dir, pattern = "*ed.txt", full.names = TRUE)

# Define color function
colfunc <- jcolors_contin(palette = "rainbow")

# Loop through each file
for (file in sfs_files) {
  
  # Extract species name (first 3 characters) and number of individuals (after 'N')
  filename <- basename(file)
  species <- substr(filename, 1, 3)
  N <- as.numeric(sub(".*N([0-9]+).*", "\\1", filename))
  allTOT <- N * 2  # Total alleles
  
  # Read data
  SFS <- read.table(file, header = TRUE)
  
  # Generate output PDF file
  pdfSFSplot <- paste0(sfs_dir, species, ".SFS.N", N, ".pdf")
  
  # Define max number of profiles
  max_profiles <- 20  
  max_lines_per_column <- (max_profiles + 2) * 4
  
  # Assign colors
  jcols <- colfunc(max_profiles + 1)
  
  # Determine number of legend columns
  colNb <- 1.0
  nb_lines <- 0
  
  for (i in 1:nrow(SFS)) {
    if (SFS[i,3] == "01") {
      SFSI <- subset(SFS, alt == SFS[i,1])
      if (length(SFSI$occ[SFSI$occ > 0]) + 1 + nb_lines > max_lines_per_column) {
        colNb <- colNb + 1.0
        nb_lines <- 0
      }
      nb_lines <- nb_lines + 1
    }
    if (SFS[i,2] > 0) {
      nb_lines <- nb_lines + 1
    }
  }
  
  # Set plot size
  ratioPointvsInch <- 0.013837
  legendfontsize <- 12.0
  charwidth <- ratioPointvsInch * legendfontsize
  nbCharperInd <- 2.0
  
  colwidth <- (N * nbCharperInd * charwidth) + (3.0 * charwidth)
  legendWidth <- 1.0 + (colNb * colwidth)
  plotWidth <- N + legendWidth
  legendlineheight <- 12.0 * ratioPointvsInch * 1.25
  plotHeight <- legendlineheight * max_lines_per_column + 2.0 * legendlineheight
  
  # Create PDF
  pdf(pdfSFSplot, width = plotWidth, height = plotHeight)
  
  # Create SFS bar plot
  p <- ggplot(data = SFS, aes(x = alt, y = occ/10000, fill = ID)) +
    ggtitle(species) +
    geom_bar(stat = "identity", colour = "black", show.legend = FALSE) +
    xlab("# alternative alleles") + ylab("Frequency") +
    scale_x_discrete(expand = c(0, 0), name = "# alternative alleles", limits = factor(c(1:allTOT))) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_bw() +
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = 'black', linewidth = 1.25),
          axis.ticks = element_line(colour = 'black', linewidth = 1.25),
          axis.ticks.length=unit(.25, "cm"),
          plot.margin = unit(c(0, legendWidth, 0, 0), "inches"),
          plot.title = element_text(colour = "black", size = 26, hjust = .5, face = "bold.italic"),
          axis.text.x = element_text(colour = "black", size = 20, angle = 0, hjust = .5, vjust = -.5, face = "plain"),
          axis.text.y = element_text(colour = "black", size = 20, angle = 0, hjust = .5, vjust = .5, face = "plain"),
          axis.title.x = element_text(colour = "black", size = 24, angle = 0, hjust = .5, vjust = .2, face = "italic"),
          axis.title.y = element_text(colour = "black", size = 24, angle = 90, hjust = .5, vjust = .5, face = "italic")) +
    scale_fill_manual(name = allTOT / 2, labels = SFS$profile[SFS$alt == allTOT / 2], values = jcols)
  
  print(p)
  
  # ---- Optimized Legend Section ----
  yinit <- plotHeight
  ypad <- legendlineheight
  ydec <- yinit
  xpad <- charwidth
  xdec <- plotWidth - (colNb * colwidth)
  column <- 0.0
  nb_lines <- 0
  matrixcolor <- rep(jcols, allTOT)
  
  # Print legend title
  # grid.text("LEGEND:", x = unit(xdec, "inches"), y = unit(yinit, "inches"), just = "left", gp = gpar(fontsize = 14))
  
  for (i in 1:nrow(SFS)) {
    currentcolor <- matrixcolor[i]
    
    if (SFS[i,3] == "01") {
      SFSI <- subset(SFS, alt == SFS[i,1])
      if (length(SFSI$occ[SFSI$occ > 0]) + 1 + nb_lines > max_lines_per_column) {
        column <- column + 1.0
        nb_lines <- 0
        ydec <- yinit
      }
      nb_lines <- nb_lines + 1
      ydec <- ydec - ypad
      grid.text(SFS[i,1], x = unit(xdec + (column * colwidth), "inches"), y = unit(ydec, "inches"), just = "left", gp = gpar(fontsize = 14))
    }
    
    if (SFS[i,2] > 0) {
      nb_lines <- nb_lines + 1
      ydec <- ydec - ypad
      
      # Add color square
      grid.polygon(x = c(xdec + (column * colwidth), xdec + (column * colwidth), 
                         xdec + (column * colwidth) + 2.0 * charwidth, 
                         xdec + (column * colwidth) + 2.0 * charwidth), 
                   y = c(ydec - (legendlineheight / 3.0), ydec + (legendlineheight / 3.0), 
                         ydec + (legendlineheight / 3.0), ydec - (legendlineheight / 3.0)), 
                   default.units = "inches", gp = gpar(fill = currentcolor))
      
      # Add text for each square
      grid.text(SFS[i,4], x = unit(xdec + (column * colwidth) + 3.0 * charwidth, "inches"), 
                y = unit(ydec, "inches"), just = "left", gp = gpar(col = "darkgrey", fontsize = legendfontsize))
    }
  }
  
  dev.off()
}

```

#### *ME analysis: haplotype trees*

```{r ME trees, echo=FALSE, message=FALSE, warning=FALSE}
library(ape)

# Define the path to the main directory
base_dir <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/MEtrees/100kbins/trees"

# List of species subfolders
species_dirs <- c("Hga", "Hti", "Ppr", "Sms", "Hrs", "Hls", "Nps", "Ass", "Aca")

# Safe tree reader that cleans NA node labels
read_safe_tree <- function(file) {
  tree <- tryCatch(read.tree(file), error = function(e) {
    message("Failed to read tree: ", file)
    return(NULL)
  })
  if (!is.null(tree) && !is.null(tree$node.label)) {
    tree$node.label[is.na(tree$node.label)] <- ""
  }
  return(tree)
}

for (sp in species_dirs) {
  message("\nProcessing species: ", sp)
  
  sp_dir <- file.path(base_dir, sp)
  if (!dir.exists(sp_dir)) {
    message("Directory does not exist: ", sp_dir, " -- skipping.")
    next
  }
  
  tree_files <- list.files(sp_dir, pattern = "\\.contree2$", full.names = TRUE)
  
  if (length(tree_files) < 2) {
    message("Not enough tree files in ", sp_dir, " -- skipping.")
    next
  }
  
  # Read and filter trees
  tree_list <- Filter(Negate(is.null), lapply(tree_files, read_safe_tree))
  
  if (length(tree_list) < 2) {
    message("Too few valid trees after filtering for ", sp, " -- skipping.")
    next
  }
  
  # Assign multiPhylo class
  class(tree_list) <- "multiPhylo"
  
  # Build majority-rule consensus tree
  consensus_tree <- consensus(tree_list, p = 0.5, check.labels = TRUE)
  
  # Write consensus tree to file
  out_file <- file.path(sp_dir, paste0(sp, "_consensus.tree"))
  write.tree(consensus_tree, file = out_file)
  
  message("Consensus tree saved to: ", out_file)
}
```

```{r MA trees, echo=FALSE, message=FALSE, warning=FALSE}
library(ape)

# Define the base directory containing species folders
base_dir <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/MEtrees/100kbins/trees"

# List of species folders
species_dirs <- c("Hga", "Hti", "Ppr", "Sms", "Hrs", "Hls", "Nps", "Ass", "Aca")

# Safe tree reading function to handle malformed node labels
read_safe_tree <- function(file) {
  tree <- tryCatch(read.tree(file), error = function(e) {
    message("Failed to read tree: ", file)
    return(NULL)
  })
  if (!is.null(tree) && !is.null(tree$node.label)) {
    tree$node.label[is.na(tree$node.label)] <- ""
  }
  return(tree)
}

# Loop over each species
for (sp in species_dirs) {
  message("\nProcessing species: ", sp)
  
  sp_dir <- file.path(base_dir, sp)
  if (!dir.exists(sp_dir)) {
    message("Directory does not exist: ", sp_dir, " -- skipping.")
    next
  }
  
  tree_files <- list.files(sp_dir, pattern = "\\.contree2$", full.names = TRUE)
  
  if (length(tree_files) < 1) {
    message("No tree files found in ", sp_dir, " -- skipping.")
    next
  }
  
  # Read all valid trees into a list
  tree_list <- Filter(Negate(is.null), lapply(tree_files, read_safe_tree))
  
  if (length(tree_list) < 1) {
    message("No valid trees after reading in ", sp, " -- skipping.")
    next
  }
  
  class(tree_list) <- "multiPhylo"
  
  # Write all trees to one file per species
  out_file <- file.path(sp_dir, paste0(sp, "_multi.trees"))
  write.tree(tree_list, file = out_file)
  
  message("Multi-tree file written to: ", out_file)
}
```

#### *ME analysis: haplotype specific dynamics*

```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(data.table)

# Define directories
bed_dir <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/haploblocks"
csv_dir <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/csv/"
output_dir <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/annotated_csv"

# Create output directory if it doesn't exist
if (!dir.exists(output_dir)) dir.create(output_dir)

# List of species prefixes based on file names
species_list <- c("Ass", "Hga", "Hls", "Hrs", "Hti", "Ppr", "Sms")

# Loop through species
for (sp in species_list) {
  
  # Define file paths
  bed_file <- file.path(bed_dir, paste0(sp, ".blocks.atleast100kb.forR.bed"))
  csv_file <- file.path(csv_dir, paste0(sp, ".var.invar.filtered.csv"))
  out_file <- file.path(output_dir, paste0(sp, ".annotated.csv"))
  
  # Read files
  large_windows <- fread(bed_file, header = FALSE)
  pop_data <- fread(csv_file)
  
  # Add column names
  colnames(large_windows) <- c("scaffold", "start_large", "end_large", "window_size", "window_name")
  
  # Convert to data.table and set keys
  setDT(large_windows)
  setDT(pop_data)
  
  # Ensure columns are in the correct type
  pop_data[, start := as.integer(start)]
  pop_data[, end := as.integer(end)]
  pop_data[, mid1 := as.integer(mid)]
  pop_data[, mid2 := as.integer(mid+2)]
  
  
  large_windows[, start_large := as.integer(start_large)]
  large_windows[, end_large := as.integer(end_large)]
  
  # Set keys for overlap
  setkey(large_windows, scaffold, start_large, end_large)
  setkey(pop_data, scaffold, mid1, mid2)
  
  # Overlap join
  annotated_data <- foverlaps(pop_data, large_windows,
                              by.x = c("scaffold", "mid1", "mid2"),
                              by.y = c("scaffold", "start_large", "end_large"),
                              type = "within", nomatch = 0)
  
  # Select and reorder columns
  # Inside your loop
  pi1_col <- paste0("pi_", sp, "1")
  pi2_col <- paste0("pi_", sp, "2")
  dxy_col <- paste0("dxy_", sp, "1_", sp, "2")
  fst_col <- paste0("Fst_", sp, "1_", sp, "2")
  
  # Subset and rename
  annotated_data <- annotated_data[, .(
    scaffold,
    start,
    end,
    mid,
    sites,
    pi1 = get(pi1_col),
    pi2 = get(pi2_col),
    dxy = get(dxy_col),
    fst = get(fst_col),
    window_size,
    window_name
  )]  
  # Write output
  fwrite(annotated_data, out_file)
  
  message("Annotated: ", sp)
}
```

```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}
# Define the input and output directories
annotated_dir <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/annotated_csv"
merged_file <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/annotated_combined.csv"

# Get list of all annotated CSVs
csv_files <- list.files(annotated_dir, pattern = "\\.annotated\\.csv$", full.names = TRUE)

# Read and combine, adding a species column inferred from filename
all_data <- rbindlist(lapply(csv_files, function(f) {
  dt <- fread(f)
  dt[, species := sub("\\.annotated\\.csv$", "", basename(f))]  # extract species name from filename
  return(dt)
}))

# Save the combined file
fwrite(all_data, merged_file)

message("Combined CSV written to: ", merged_file)
```


```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}
# Mean diversity stats per window (across species)
summary_stats <- all_data[, .(
  mean_pi1 = mean(pi1, na.rm = TRUE),
  mean_pi2 = mean(pi2, na.rm = TRUE),
  mean_dxy = mean(dxy, na.rm = TRUE),
  mean_fst = mean(fst, na.rm = TRUE)
), by = .(species, scaffold, window_name, window_size)]

head(summary_stats)


```

```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)

# Assuming all_data is already loaded (merged annotated CSV)
# If not, read it:
all_data <- fread("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/annotated_combined.csv")

# Ensure data.table
setDT(all_data)

# First, calculate unique window ranks
window_ranks <- unique(all_data[, .(species, scaffold, window_name, window_size)])
window_ranks[, window_rank := frank(-window_size, ties.method = "min"), by = .(species, scaffold)]

# Merge ranks back into the full dataset
all_data <- merge(all_data, window_ranks[, .(species, scaffold, window_name, window_rank)],
                  by = c("species", "scaffold", "window_name"), all.x = TRUE)


# reformat
plot_data <- reshape2::melt(all_data, id=c("species", "scaffold", "window_name", "start", "end", "mid", "sites", "window_size", "window_rank"))

plot_data$stat <- gsub("[^A-Za-z]", "", plot_data$variable)

# Violin plot: Pi
dt <- plot_data %>% filter(plot_data$stat =="pi" &
                             plot_data$window_rank < 5 &
                             plot_data$window_size > 1e6 &
                             plot_data$species == "Ppr")

dt$window_rank <- as_factor(dt$window_rank)

# pi <- ggplot(dt,aes(variable, log10(value)))+  theme_bw()+
#   geom_quasirandom(aes(color=variable, fill = variable), alpha=.3,size=.1, dodge.width = 0.9, width = 0.2)+
#   geom_boxplot(aes(color= variable, fill = variable), color = "black", width=0.1,alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
#   geom_violin(aes(color= variable, fill = variable), alpha=.4, linewidth=.3, scale = "width") +
#   theme(panel.border = element_blank(), 
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(), 
#         axis.line = element_line(colour = "black")) +
#   stat_compare_means(comparisons = list(c("pi1", "pi2")), aes(label = ..p.signif..), label.x = 3, label.y = -2)+
#   facet_grid(species~scaffold+window_rank,  scales = "free", space = "free_x", switch = "x") +
#   scale_fill_manual(name="haplotype",values = c(
#     "pi1"="#F56B39",
#     "pi2"="#979797"))+
#   scale_color_manual(name="haplotype",values = c(
#     "pi1"="#F56B39",
#     "pi2"="#979797"))+
#   # # coord_cartesian(ylim = c(-6,-3))+
#   scale_y_continuous(name=expression(paste("log"[10]," nucleotide diversity")), breaks = c(-3,-2,-1))+
#   theme(legend.position = "top")+xlab(NULL)+
#   theme(panel.spacing.x=unit(-.1,"lines"),
#         panel.spacing.y=unit(.5,"lines"),
#         panel.background = element_blank(),
#         axis.text.x = element_blank(), #element_text(size=9),
#         axis.ticks.x = element_blank(),
#         panel.spacing = unit(.1, "cm"),
#         strip.background=element_rect(fill='white', colour='black'),
#         axis.text.y=element_text(size=10),
#         plot.title = element_text(size = 10)
#   ) #+ ggtitle("A")
# 
# 
# pi

```

```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)

# Directory containing the annotated CSVs
input_dir <- "~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/csv/"
csv_files <- list.files(input_dir, pattern = "\\.csv$", full.names = TRUE)

# Function to load and standardize each file
load_and_standardize <- function(f) {
  # Extract species from filename
  sp <- sub(".var.invar.filtered.csv$", "", basename(f))
  
  # Read data
  dt <- fread(f)
  
  # Dynamically identify original column names
  pi1_col <- paste0("pi_", sp, "1")
  pi2_col <- paste0("pi_", sp, "2")
  dxy_col <- paste0("dxy_", sp, "1_", sp, "2")
  fst_col <- paste0("Fst_", sp, "1_", sp, "2")
  
  # Rename relevant columns for merging
  setnames(dt, c(pi1_col, pi2_col, dxy_col, fst_col),
           c("pi1",    "pi2",    "dxy",    "fst"), skip_absent = TRUE)
  
  # Add species column
  dt[, species := sp]
  
  return(dt)
}

# Load and combine all
all_data <- rbindlist(lapply(csv_files, load_and_standardize), fill = TRUE)

# Preview
head(all_data)

```

```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}

# Violin plot: Pi

# reformat
plot_data <- reshape2::melt(all_data, id=c("scaffold", "start", "end", "mid", "sites", "species"))

plot_data$stat <- gsub("[^A-Za-z]", "", plot_data$variable)


# plot delta Pi for each species:

all_data$deltaPi <- abs(all_data$pi1 - all_data$pi2)

# extract species of interest:
all_data <- all_data %>% filter(species != "Ona"|
                                  species != "Osa"|
                                  species != "Nps"|
                                  species != "Aca")
# reproductive mode:
all_data <- mutate(all_data, reproMode = ifelse(all_data$species == "Ppr" |
                                                  all_data$species == "Hrs" |
                                                  all_data$species == "Hls" |
                                                  all_data$species == "Ass", "asexual", "sexual"))

all_data$species <- factor(all_data$species, levels = c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))


# Add full species name:
species_lookup <- c(
  "Hrs" = "H. rufulus",
  "Sms" = "S. magnus",
  "Ass" = "A. striculus",
  "Hga" = "H. gibba",
  "Ppr" = "P. peltifer",
  "Hti" = "H. thienemanni",
  "Hls" = "H. lacustris"
)


# Map the species names
all_data[, Species := species_lookup[species]]


# Set the order of the Species factor
all_data[, Species := factor(Species, levels = rev(c(
  "H. rufulus", "S. magnus", "A. striculus", "H. gibba",
  "P. peltifer", "H. thienemanni", "H. lacustris"
)))]


# Plot:
dpi <- ggplot(all_data,aes(species, log10(deltaPi)))+  theme_bw()+
  # geom_quasirandom(aes(color=reproMode, fill = reproMode), alpha=.3,size=.1, dodge.width = 0.9, width = 0.2)+
  geom_boxplot(aes(color= reproMode, fill = reproMode), color = "black", width=0.1,alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  geom_violin(aes(color= reproMode, fill = reproMode), alpha=.4, linewidth=.3, scale = "width") +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_y_continuous(name=expression(paste("log"[10]," delta nuc. diversity")))+
  theme(legend.position = "none")+xlab(NULL)+
  theme(#panel.spacing.x=unit(-.1,"lines"),
        #panel.spacing.y=unit(.5,"lines"),
        # panel.background = element_blank(),
        # axis.text.x = element_text(angle = 45, hjust = 1),
        # axis.ticks.x = element_blank(),
        # panel.spacing = unit(.1, "cm"),
        # strip.background=element_rect(fill='white', colour='black'),
        # axis.text.y=element_text(size=10),
        # plot.title = element_text(size = 10)
  ) #+ ggtitle("A")


dpi

# ggsave("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/DeltaPi_Boxplot.pdf", dpi, width = 3, height = 2.2)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/DeltaPi_Boxplot.pdf", dpi, width = 2.6, height = 2.2)


# Make stats:

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(deltaPi ~ Species, data = all_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(all_data$deltaPi, all_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load thsi
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

# Dxy
# Plot:
dxy <- ggplot(all_data,aes(species, log10(dxy)))+  theme_bw()+
  # geom_quasirandom(aes(color=reproMode, fill = reproMode), alpha=.3,size=.1, dodge.width = 0.9, width = 0.2)+
  geom_boxplot(aes(color= reproMode, fill = reproMode), color = "black", width=0.1,alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  geom_violin(aes(color= reproMode, fill = reproMode), alpha=.4, linewidth=.3, scale = "width") +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_y_continuous(name=expression(paste("log"[10]," ", italic(d)[xy])), breaks=c(-4,-3,-2))+
  theme_classic2() +
  theme(legend.position = "top")+xlab(NULL)+
  theme(#panel.spacing.x=unit(-.1,"lines"),
        # panel.spacing.y=unit(.5,"lines"),
        # panel.background = element_blank(),
        # axis.text.x = element_text(angle = 45, hjust = 1),
        # axis.ticks.x = element_blank(),
        # panel.spacing = unit(.1, "cm"),
        # strip.background=element_rect(fill='white', colour='black'),
        # axis.text.y=element_text(size=10),
        # plot.title = element_text(size = 10)
  ) #+ ggtitle("A")

dxy

# ggsave("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/dxy_Boxplot.pdf", dxy, width = 3, height = 2.2)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/dxy_Boxplot.pdf", dxy, width = 3, height = 2.5)

# Make stats:

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(dxy ~ Species, data = all_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(all_data$dxy, all_data$Species, p.adjust.method = "BH")

print(pwc)

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)


myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters


# Fst
all_data[all_data$fst <0,]$fst <- 0

# Plot:
fst <- ggplot(all_data,aes(species, (fst)))+  theme_bw()+
  # geom_quasirandom(aes(color=reproMode, fill = reproMode), alpha=.3,size=.1, dodge.width = 0.9, width = 0.2)+
  geom_boxplot(aes(color= reproMode, fill = reproMode), color = "black", width=0.1,alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  geom_violin(aes(color= reproMode, fill = reproMode), alpha=.4, linewidth=.3, scale = "width") +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  ylab(expression(italic("F")[ST]))+
  theme(legend.position = "none")+xlab(NULL)+
  theme(panel.spacing.x=unit(-.1,"lines"),
        panel.spacing.y=unit(.5,"lines"),
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        # axis.ticks.x = element_blank(),
        panel.spacing = unit(.1, "cm"),
        strip.background=element_rect(fill='white', colour='black'),
        axis.text.y=element_text(size=10),
        plot.title = element_text(size = 10)
  ) #+ ggtitle("A")


fst

ggsave("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/diversity/Fst_Boxplot.pdf", fst, width = 3, height = 2.2)

# Make stats:

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(fst ~ Species, data = all_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(all_data$fst, all_data$Species, p.adjust.method = "BH")

print(pwc)

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)


myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

```

```{r popGen: general, echo=FALSE, message=FALSE, warning=FALSE}
# plot heterozygosity:

# Load required libraries
library(ggplot2)
library(dplyr)

# Set working directory (change this to your folder)
path = "~/sciebo/Postdoc/sexAsex/popGen/diversity/het/"

# List all .het.txt files in the directory
files<-dir(path,pattern =".txt")

# Initialize an empty dataframe to store results
heterozygosity_data <- data.frame()

# Loop through each file and read data
for (file in files) {
  # Extract species name from the filename (assuming format: Species.het.txt)
  species_name <- gsub(".het.txt", "", file)
  
  # Read data
  df <- read.table(paste(path, file, sep = ""), header = TRUE, sep = "\t")
  
  # Add species column
  df$Species <- species_name
  
  # Combine data
  heterozygosity_data <- rbind(heterozygosity_data, df)
}

# Convert species to a factor
heterozygosity_data$Species <- as.factor(heterozygosity_data$Species)

heterozygosity_data <- heterozygosity_data %>% filter(Species != "Ona"|
                                                        Species != "Osa"|
                                                        Species != "Nps"|
                                                        Species != "Aca")


# add reproductive mode
heterozygosity_data <- mutate(heterozygosity_data, reproMode = ifelse(heterozygosity_data$Species == "Ppr" |
                                                                        heterozygosity_data$Species == "Hrs" |
                                                                        heterozygosity_data$Species == "Hls" |
                                                                        heterozygosity_data$Species == "Ass", "asexual", "sexual"))

# order species:
heterozygosity_data$Species <- factor(heterozygosity_data$Species, levels = rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls")))


# Plot boxplot of heterozygosity rate
p <- ggplot(heterozygosity_data, aes(x = Species, y = Heterozygosity_Rate*100, fill = Species)) +
  geom_violin(aes(fill = reproMode, color = reproMode), alpha=.5, linewidth=.3,  drop = FALSE, scale = "width") +
  geom_boxplot(aes(fill = reproMode), color = "black", width=0.1, alpha=.8, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  theme_classic2() +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_y_continuous(breaks = c(.5,1))+
  labs(title = NULL,
       x = NULL,
       y = "SNP heterozygosity") +
  theme(#axis.text.x = element_text(angle = 90, hjust = 1),
        # panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "none") + coord_flip()


p

# Save plot
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/Heterozygosity_Boxplot.pdf", p, width = 2.2, height = 2.5)


# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(Heterozygosity_Rate ~ Species, data = heterozygosity_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(heterozygosity_data$Heterozygosity_Rate, heterozygosity_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load thsi
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

```

```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}
# plot nucleotide diversity:
## Distribution:

# Set working directory (change this to your folder)
path = "~/sciebo/Postdoc/sexAsex/popGen/diversity/div/csv/"

# List all .het.txt files in the directory
files<-dir(path,pattern =".csv")

# Initialize an empty dataframe to store results
diversity_data <- data.frame()

# Loop through each file and read data
for (file in files) {
  # Extract species name from the filename (assuming format: Species.var.invar.filtered.csv)
  species_name <- gsub(".var.invar.filtered.csv", "", file)
  
  # Read data
  df <- read.table(paste(path, file, sep = ""), header = TRUE, sep = ",")
  names(df)[names(df) == paste("pi_",species_name, sep = "")] <- "pi"
  
  # Add species column
  df$Species <- species_name
  
  # Combine data
  diversity_data <- rbind(diversity_data, df)
}

# Convert species to a factor
diversity_data$Species <- as.factor(diversity_data$Species)

# add reproductive mode
diversity_data <- mutate(diversity_data, reproMode = ifelse(diversity_data$Species == "Ppr" |
                                                              # diversity_data$Species == "Nps" |
                                                              # diversity_data$Species == "Ona" |
                                                              diversity_data$Species == "Hrs" |
                                                              diversity_data$Species == "Hls" |
                                                              diversity_data$Species == "Ass", "asexual", "sexual"))

# order species:
diversity_data$Species <- factor(diversity_data$Species, levels = rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls")))

# Plot boxplot of heterozygosity rate
p <- ggplot(diversity_data, aes(x = Species, y = log10(pi), fill = Species)) +
  geom_violin(aes(fill = reproMode, color = reproMode), alpha=.5, linewidth=.3,  drop = FALSE, scale = "width") +
  geom_boxplot(aes(fill = reproMode), color = "black", width=0.1,alpha=.8, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  theme_classic2() +
  labs(title = NULL,
       x = NULL,
       y = expression(paste("log"[10], " nucleotide diversity"))) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
        #panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "none") + coord_flip()


p

# Save plot
# ggsave("~/sciebo/Postdoc/sexAsex/popGen/diversity/div/Diversity_Boxplot.pdf", p, width = 3, height = 2.2)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/Diversity_Boxplot.pdf", p, width = 2.2, height = 2.5)

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(pi ~ Species, data = diversity_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(diversity_data$pi, diversity_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load thsi
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

## Genome-wide distribution:

# Extract numeric part of scaffold names
diversity_data$CHR <- as.numeric(gsub("\\D", "", diversity_data$scaffold))

# Filter scaffolds larger than 1Mb
large_scaffolds <- diversity_data %>%
  group_by(Species, scaffold) %>%
  filter((max(mid) - min(mid)) >= 2e6) %>%
  ungroup()

# Generate short names for scaffolds
short <- unique(large_scaffolds$CHR)
names(short) <- unique(large_scaffolds$CHR)

# Get unique species
large_scaffolds$Species <- factor(large_scaffolds$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  
species_list <- levels(large_scaffolds$Species)

# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(sp) {
  species_data <- large_scaffolds %>% filter(Species == sp)
  
  ggplot(species_data, aes(x = mid, y = log10(pi))) +
    geom_jitter(aes(color=reproMode), size=.3, alpha=.2) +
    geom_smooth(aes(color=reproMode, fill = reproMode), method = "loess", span = .1, se=T, size=.4, alpha=.2) +
    facet_grid(cols = vars(CHR), labeller = labeller(scaffold=short),
               scales = 'free_x', space = 'free_x', switch = 'x') +  # Arrange scaffolds in rows
    theme_classic() +
    scale_fill_manual(name="reproMode",values = c(
      "sexual"="#F56B39",
      "asexual"="#11A7BB"))+
    scale_color_manual(name="reproMode",values = c(
      "sexual"="#F56B39",
      "asexual"="#11A7BB"))+
    ggtitle(sp) +  # Title for each species
    theme(panel.spacing = unit(.1, "cm"),
          panel.background = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom") +
    coord_cartesian(ylim = c(-4, -1)) +
    scale_y_continuous(breaks = c(-4, -3, -2, -1), labels = c(-4, -3, -2, -1)) +
    scale_x_continuous(expand = c(0, 0)) +
    xlab(NULL) +
    ylab(expression(paste("nuc. diversity")))
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 1, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "bottom")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/diversity/div/Diversity_distribution.w25k10ksites.pdf",width=16,height=12)


```

```{r popGen: haplotypic, echo=FALSE, message=FALSE, warning=FALSE}
#############################
# plot Tajima's D:
## Distribution:

# Set working directory (change this to your folder)
path = "~/sciebo/Postdoc/sexAsex/popGen/diversity/tajD/"

# List all .het.txt files in the directory
files<-dir(path,pattern =".Tajima.D")

# Initialize an empty dataframe to store results
tajima_data <- data.frame()

# Loop through each file and read data
for (file in files) {
  # Extract species name from the filename (assuming format: Species.var.invar.filtered.tajD.Tajima.D)
  species_name <- gsub(".var.invar.filtered.tajD.Tajima.D", "", file)
  
  # Read data
  df <- read.table(paste(path, file, sep = ""), header = TRUE, sep = "\t")
  names(df)[names(df) == paste("pi_",species_name, sep = "")] <- "Tajima"
  
  # Add species column
  df$Species <- species_name
  
  # Combine data
  tajima_data <- rbind(tajima_data, df)
}

# Convert species to a factor
tajima_data$Species <- as.factor(tajima_data$Species)

# add reproductive mode
tajima_data <- mutate(tajima_data, reproMode = ifelse(tajima_data$Species == "Ppr" |
                                                        # tajima_data$Species == "Nps" |
                                                        # tajima_data$Species == "Ona" |
                                                        tajima_data$Species == "Hrs" |
                                                        tajima_data$Species == "Hls" |
                                                        tajima_data$Species == "Ass", "asexual", "sexual"))

# order species:
tajima_data$Species <- factor(tajima_data$Species, levels = rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))) 

# Plot boxplot of heterozygosity rate
p <- ggplot(tajima_data, aes(x = Species, y = TajimaD, fill = Species)) +
  geom_violin(aes(fill = reproMode, color = reproMode), alpha=.5, linewidth=.3,  drop = FALSE, scale = "width") +
  geom_boxplot(aes(fill = reproMode), color = "black", width=0.1,alpha=.8, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  theme_classic2() +
  labs(title = NULL,
       x = NULL,
       y = expression(paste("Tajima's ",italic(D)))) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
        #panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "none") + coord_flip()


p

# Save plot
# ggsave("~/sciebo/Postdoc/sexAsex/popGen/diversity/tajD/tajima_Boxplot.pdf", p, width = 3, height = 2.2)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tajima_Boxplot.pdf", p, width = 2.2, height = 2.5)

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(TajimaD ~ Species, data = tajima_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(tajima_data$TajimaD, tajima_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load thsi
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

## Genome-wide distribution:

# Extract numeric part of scaffold names
tajima_data$CHR <- as.numeric(gsub("\\D", "", tajima_data$CHROM))

# Filter scaffolds larger than 1Mb
large_scaffolds <- tajima_data %>%
  group_by(Species, CHROM) %>%
  filter((max(BIN_START) - min(BIN_START)) >= 2e6) %>%
  ungroup()

# Generate short names for scaffolds
short <- unique(large_scaffolds$CHR)
names(short) <- unique(large_scaffolds$CHR)

# Get unique species
large_scaffolds$Species <- factor(large_scaffolds$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  
species_list <- levels(large_scaffolds$Species)

# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(sp) {
  species_data <- large_scaffolds %>% filter(Species == sp)
  
  ggplot(species_data, aes(x = BIN_START, y = (TajimaD))) +
    geom_jitter(aes(color=reproMode), size=.3, alpha=.2) +
    geom_smooth(aes(color=reproMode, fill = reproMode), method = "loess", span = .1, se=T, size=.4, alpha=.2) +
    facet_grid(cols = vars(CHR), labeller = labeller(scaffold=short),
               scales = 'free_x', space = 'free_x', switch = 'x') +  # Arrange scaffolds in rows
    theme_classic() +
    scale_fill_manual(name="reproMode",values = c(
      "sexual"="#F56B39",
      "asexual"="#11A7BB"))+
    scale_color_manual(name="reproMode",values = c(
      "sexual"="#F56B39",
      "asexual"="#11A7BB"))+
    ggtitle(sp) +  # Title for each species
    theme(panel.spacing = unit(.1, "cm"),
          panel.background = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom") +
    # coord_cartesian(ylim = c(-4, -1)) +
    scale_y_continuous(breaks = c(-3, -2, -1, 1, 2, 3), labels = c(-3, -2, -1, 1, 2, 3)) +
    scale_x_continuous(expand = c(0, 0)) +
    xlab(NULL) +
    ylab(expression(paste("Tajima's ", italic(D))))
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 1, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "bottom")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/diversity/tajD/Tajima_distribution.w25k.pdf",width=16,height=16)

```

```{r Hyphy, echo=FALSE, message=FALSE, warning=FALSE}
# fitmg94
fitmg94 <- read_delim("~/sciebo/Postdoc/sexAsex/popGen/hyphy/FitMG94.100perc.scOGs.summary.tsv")

# add gene/species columns:
fitmg94[c("geneID", "sp")] <- str_split_fixed(fitmg94$Gene, '_', 4)[,c(1,3)]

# add haplotype
dt <- fitmg94 %>% filter(!grepl(glob2rx('Node*'), ignore.case = T,fitmg94$geneID))

dt <- mutate(dt, haplotype = ifelse(grepl(glob2rx('*_1'), ignore.case = F,dt$Gene), "A","B"))

# order spcies
dt$Species <- factor(dt$sp, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  


#
# make Paired line plots
good <- dt[!dt$OG %in% dt[dt$MLE > 1,]$OG,]


# # make plots:
# ggplot(good,aes(Species, MLE))+  theme_bw()+
#   geom_quasirandom(aes(color=haplotype, fill = haplotype), alpha=.5,size=.1, dodge.width = 0.9, width = 0.2)+
#   geom_boxplot(aes(color= haplotype, fill = haplotype), color = "black", width=0.1,alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
#   geom_violin(aes(color= haplotype, fill = haplotype), alpha=.4, linewidth=.3, scale = "width", drop = FALSE) +
#   theme(panel.border = element_blank(), 
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(), 
#         axis.line = element_line(colour = "black")) +
#   scale_y_continuous(name=expression(paste("dN/dS")))+
#   theme(legend.position = "top")+xlab(NULL)+
#   theme(panel.spacing.x=unit(-.1,"lines"),
#         panel.spacing.y=unit(.5,"lines"),
#         panel.background = element_blank(),
#         axis.text.x = element_text(size=9),
#         panel.spacing = unit(.1, "cm"),
#         strip.background=element_rect(fill='white', colour='black'),
#         axis.text.y=element_text(size=10),
#         plot.title = element_text(size = 10)
#   ) 


# add directionality:
# Ensure haplotypes are labeled as "A" and "B" consistently
df <- good 

# Compare MLE values for haplotypes A and B within each geneID and population
df_comparison <- df %>%
  dplyr::group_by(geneID, Species) %>%
  dplyr::reframe(
    geneID = first(geneID),
    population = first(Species),
    MLE_A = MLE[haplotype == "A"],
    MLE_B = MLE[haplotype == "B"]
  ) %>%
  dplyr::mutate(
     IntensePurOn = case_when(
      MLE_A < MLE_B ~ "A",
      MLE_A > MLE_B ~ "B",
      TRUE ~ "same"
    )
  )


# write.table(df_comparison, "~/sciebo/Postdoc/sexAsex/popGen/hyphy/FitMG94_df_comparison_final.tsv",
#             sep = "\t",col.names = T, quote = F, row.names = F, na = "NA")


# calculate difference between A and B:
df_comparison$difference <- abs(df_comparison$MLE_A - df_comparison$MLE_B)


# only different
df_final <- reshape2::melt(df_comparison, id=c("geneID", "population","Species", "difference", "IntensePurOn")) #[df_comparison$IntensePurOn != "same",]


df_final$allele = gsub("MLE_", "", df_final$variable)

df_final <- mutate(df_final, reproMode = ifelse(df_final$Species == "Ppr" |
                                                # df_final$Species == "Nps" |
                                                # df_final$Species == "Ona" |
                                                df_final$Species == "Hrs" |
                                                df_final$Species == "Hls" |
                                                df_final$Species == "Ass", "asexual", "sexual"))



# # plot:
# f <- ggplot(df_final, aes(x = allele, y = value)) +
#   stat_summary(
#     aes(group = geneID, color = IntensePurOn),
#     fun = mean, geom = "path", alpha = 0.1, linewidth = .4
#   ) +
#   geom_quasirandom(aes(color = variable), alpha=.5,size=.5, dodge.width = 0.9, width = 0.2)+
#   # geom_point(aes(color = variable), alpha = .2, size = 0.5) +
#   scale_x_discrete("") +
#   scale_y_continuous(breaks = c(0.2, 0.4, 0.6, 0.8, 1.0)) +  # <-- Add this line
#   scale_color_manual(values=c("#003B73","#BFD7ED", "grey30", "#FFA405", "#59981A")) +
#   facet_grid(~population, scales = "free", space = "free_x", switch = "x") +
#   theme_classic2() +
#   theme(#panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(colour = "black"),
#         axis.ticks = element_line(color = "black", linewidth = 0.3),
#         axis.ticks.length = unit(0.2, "cm"),
#         axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
#         axis.title.x = element_blank(),
#         legend.title = element_text(size = 10),
#         legend.text = element_text(size = 10),
#         legend.position = "none"
#   ) +
#   ylab("omega") +
#   xlab("allele")

# ggsave("~/sciebo/Postdoc/sexAsex/popGen/hyphy/FitMG94_hapA_vs_hapB.pdf",plot = f,width=5,height=2.5)
# ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/FitMG94_hapA_vs_hapB.png",plot = f,width=5,height=3, dpi = 1500)



# Stats
su <- group_by(df_final, population, IntensePurOn) %>%
  dplyr::summarise(
    count = length(unique(geneID))
  )

su

# write_delim(su, "~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/FitMG94_hapA_vs_hapB_counts.txt", delim = "\t")


# plot:
dif <- df_final %>% filter(variable == "MLE_A")

# ff <- ggplot(dif, aes(x = population, y = -log10(difference))) +
#   geom_quasirandom(aes(colour = reproMode), alpha=.3,size=.7, dodge.width = 0.9, width = 0.2)+
#   geom_boxplot(aes(fill = reproMode), color = "black", width=0.1,alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
#   scale_color_manual(
#     name = "reproMode",
#     values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")
#     ) +
#   scale_fill_manual(
#     name = "reproMode",
#     values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")
#   ) +
#   theme_classic2() +
#   theme(#panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
#         # panel.grid.major = element_blank(),
#         # panel.grid.minor = element_blank(),
#         # axis.line = element_line(colour = "black"),
#         # axis.ticks = element_line(color = "black", linewidth = 0.3),
#         # axis.ticks.length = unit(0.2, "cm"),
#         # axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
#         # axis.title.x = element_blank(),
#         # legend.title = element_text(size = 10),
#         # legend.text = element_text(size = 10),
#         legend.position = "top"
#   ) +
#   ylab(expression(paste(Delta," ", omega, ("A~B")))) +
#   xlab(NULL)


# ggsave("~/sciebo/Postdoc/sexAsex/popGen/hyphy/diff_FitMG94_hapA_vs_hapB.pdf",plot = ff,width=3.5,height=2.5)
# ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/diff_FitMG94_hapA_vs_hapB.pdf",plot = ff,width=3,height=2.5)

# stats:
su2 <- dif %>%
  group_by(population) %>%
  summarise(
    count_gt1 = sum(difference > 0, na.rm = TRUE),
    meanDif   = mean(difference, na.rm = TRUE),
    medianDif = median(difference, na.rm = TRUE),
    .groups = "drop"
  )

su2

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(difference ~ population, data = dif)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(dif$difference, dif$population, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load this
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

```


```{r Phasing quality, echo=FALSE, message=FALSE, warning=FALSE}
# Load required libraries
library(ggplot2)
library(data.table)
library(ggpubr)
library(dplyr)

# Define folder with .bed files
bed_folder <- "/Users/mohammederrbii/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/haploblocks"

# List .bed files
bed_files <- list.files(path = bed_folder, pattern = "atleast100kb.bed$", full.names = TRUE)

# Read and combine all BED files
haplotype_data <- do.call(rbind, lapply(bed_files, function(file) {
  df <- fread(file, header = FALSE)
  species <- sub("\\..*$", "", basename(file))  # Extract species from filename
  df <- df[, .(Species = species, length = V4)]
  return(df)
}))

# Add reproductive mode using your logic
haplotype_data <- haplotype_data %>%
  mutate(reproMode = ifelse(Species %in% c("Ppr", "Hrs", "Hls", "Ass"), 
                            "asexual", "sexual"))

# Set factor levels for species (for plot order)
haplotype_data$Species <- factor(haplotype_data$Species, levels = c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))

# Plot
p <- ggplot(haplotype_data, aes(x = Species, y = log10(length / 1e5), fill = Species)) +
  geom_violin(aes(fill = reproMode, color = reproMode), alpha = .5, linewidth = .3, scale = "width") +
  geom_boxplot(aes(fill = reproMode), color = "black", width = 0.1, alpha = .8,
               linewidth = .3, position = position_dodge(width = 0.9), outlier.shape = NA) +
  theme_classic2() +
  labs(title = NULL,
       x = NULL,
       y = expression(paste("haploblocks length (kb)"))) +
  scale_fill_manual(name = "reproMode", values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")) +
  scale_color_manual(name = "reproMode", values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "none")

# Show plot
print(p)

# Save plot
ggsave("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/haploblocks/haploblocks_Boxplot.pdf", p, width = 3, height = 2.2)

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(length ~ Species, data = haplotype_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(haplotype_data$length, haplotype_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load thsi
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

```

#### *SV analysis*

```{r Sniffles, echo=FALSE, message=FALSE, warning=FALSE}

# List all .table files in the directory
table_files <- list.files(path = "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2", pattern = "\\.sniffles2_variants\\.table$", full.names = TRUE)
species_list <- sub("\\.sniffles2_variants\\.table$", "", basename(table_files))

# Initialize list to store data
data_list <- list()

# Loop over species
for (species in species_list) {
  cat("Processing:", species, "\n")
  
  table_file <- file.path("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2", paste0(species, ".sniffles2_variants.table"))
  
  if (!file.exists(table_file)) {
    warning("Missing table file: ", table_file)
    next
  }
  
  # Read the data and assign to list using species name as key
  data_list[[species]] <- read.table(table_file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
  data_list[[species]]$DP <- rowSums(data_list[[species]][, 13:ncol(data_list[[species]])])
  
  data_list[[species]] <- data_list[[species]][, c(1:12,ncol(data_list[[species]]))]
  
  data_list[[species]]$Species <- species

}

# Combine all tables into a single data frame by row binding
sfinal_data_list <- bind_rows(data_list)

# set colors
cols <-c(
  "AF"="#ff6666")

# Get unique species
species_list <- unique(sfinal_data_list$Species)

# AF
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- sfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = AF, after_stat(density), fill = "AF"), binwidth = 0.05, alpha = 0.5)+
    labs(title = "Allele frequency", x = "Frequency", y = "Density")+
    scale_fill_manual(name="Allele",values=cols)+ 
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
    ) + ggtitle(species)# Title for each species
  
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/Freq.pdf",width=9,height=6)

#MAF
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- sfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = MAF, after_stat(density), fill = "MAF"), binwidth = 0.05, alpha = 0.5)+
    labs(title = "Allele frequency", x = "Frequency", y = "Density")+
    scale_fill_manual(name="Allele",values="green")+ 
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
    ) + ggtitle(species)# Title for each species
  
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/MAF.pdf",width=9,height=6)

#DP
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- sfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = DP/NS, stat(density)), binwidth = 1, fill = "grey")+ 
    labs(title = "Mean depth per site")+
    xlim(c(0,150))+
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species, ": ", round(quantile(species_data$DP/species_data$NS, 0.05, na.rm = T),digits = 0), " - ", round(quantile(species_data$DP/species_data$NS, 0.95, na.rm = T), 0))) # Title for each species 
  
})


# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

# dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/Depth.pdf",width=9,height=6)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/Depth.ave.pdf",width=9,height=6)

#F_MISSING
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- sfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = F_MISSING, stat(density)),  binwidth = 0.1, fill = "grey")+
    labs(title = "Missingness on a per-site basis", x = "Frequency (binwidth = 0.1)", y = "Density")+
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species)) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/Miss.pdf",width=9,height=6)

#QUAL
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- sfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = abs(QUAL), stat(density)),  binwidth = 1, fill = "grey")+
    labs(title = "Missingness on a per-site basis", x = "Frequency (binwidth = 0.1)", y = "Density")+
    theme_bw() + 
    # xlim(c(0,10))+
    theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species)) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/QUAL.pdf",width=9,height=6)

#SV LENTH
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- sfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = abs(SVLEN)/1000, stat(density)),  binwidth = 0.1, fill = "grey")+
    labs(title = "Missingness on a per-site basis", x = "Frequency (binwidth = 0.1)", y = "Density")+
    theme_bw() + 
    xlim(c(0,10))+
    theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species)) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/SVLEN.lim.pdf",width=9,height=6)

write_delim(sfinal_data_list, "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sfinal_data_list.txt", delim = "\t")

```


```{r GraffiTE, echo=FALSE, message=FALSE, warning=FALSE}

# List all .table files in the directory
table_files <- list.files(path = "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/", pattern = "\\.GraffiTE.merged.genotypes.annotated.biall\\.table$", full.names = TRUE)
species_list <- sub("\\.GraffiTE.merged.genotypes.annotated.biall\\.table$", "", basename(table_files))

# Initialize list to store data
data_list <- list()

# Loop over species
for (species in species_list) {
  cat("Processing:", species, "\n")
  
  table_file <- file.path("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/", paste0(species, ".GraffiTE.merged.genotypes.annotated.biall.table"))
  
  if (!file.exists(table_file)) {
    warning("Missing table file: ", table_file)
    next
  }
  
  # Read the data and assign to list using species name as key
  data_list[[species]] <- read.table(table_file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
  # data_list[[species]]$DP <- rowSums(data_list[[species]][, 13:ncol(data_list[[species]])])
  
  # data_list[[species]] <- data_list[[species]][, c(1:12,ncol(data_list[[species]]))]
  
  data_list[[species]]$Species <- species

}

# Combine all tables into a single data frame by row binding
gfinal_data_list <- bind_rows(data_list)

# set colors
cols <-c(
  "AF"="#ff6666")

# Get unique species
species_list <- unique(gfinal_data_list$Species)

# AF
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- gfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = as.numeric(as.character(AF)), after_stat(density), fill = "AF"), binwidth = 0.05, alpha = 0.5)+
    labs(title = "Allele frequency", x = "Frequency", y = "Density")+
    scale_fill_manual(name="Allele",values=cols)+ 
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
    ) + ggtitle(species)# Title for each species
  
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/Freq.pdf",width=9,height=6)

#MAF
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- gfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = as.numeric(as.character(MAF)), after_stat(density), fill = "MAF"), binwidth = 0.05, alpha = 0.5)+
    labs(title = "Allele frequency", x = "Frequency", y = "Density")+
    scale_fill_manual(name="Allele",values="green")+ 
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
    ) + ggtitle(species)# Title for each species
  
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/MAF.pdf",width=9,height=6)

#DP
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- gfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = as.numeric(as.character(DP/NS)), stat(density)), binwidth = 1, fill = "grey")+ 
    labs(title = "Mean depth per site")+
    xlim(c(0,200))+
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species, ": ", round(quantile(species_data$DP/species_data$NS, 0.05, na.rm = T),digits = 0), " - ", round(quantile(species_data$DP/species_data$NS, 0.95, na.rm = T), 0))) # Title for each species 
  
})


# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/Depth.pdf",width=9,height=6)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/Depth.ave.pdf",width=9,height=6)

#F_MISSING
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- gfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = as.numeric(as.character(F_MISSING)), stat(density)),  binwidth = 0.1, fill = "grey")+
    labs(title = "Missingness on a per-site basis", x = "Frequency (binwidth = 0.1)", y = "Density")+
    theme_bw() + theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species)) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/Miss.pdf",width=9,height=6)


#QUAL
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- gfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = abs(as.numeric(as.character(QUAL))), stat(density)),  binwidth = 100, fill = "grey")+
    labs(title = "Missingness on a per-site basis", x = "Frequency (binwidth = 100)", y = "Density")+
    theme_bw() + 
    xlim(c(0,10000))+
    theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species)) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/QUAL.pdf",width=9,height=6)


# SVLEN
# Generate plots for each species and store them in a list
plot_list <- lapply(species_list, function(species) {
  species_data <- gfinal_data_list %>% filter(Species == species)
  
  ggplot(species_data) + 
    geom_histogram(aes(x = abs(as.numeric(as.character(SVLEN)))/1000, stat(density)),  binwidth = 0.1, fill = "grey")+
    labs(title = "Missingness on a per-site basis", x = "Frequency (binwidth = 0.1)", y = "Density")+
    theme_bw() + 
    xlim(c(0,10))+
    theme(
      plot.title = element_text(size=8),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      axis.text=element_text(size=8),
      axis.title=element_text(size=8,face="bold"),
      panel.border = element_blank(), panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")
    ) + ggtitle(paste0(species)) # Title for each species 
  
})

# Combine all species plots into one figure and collect a single legend
final_plot <- wrap_plots(plot_list) + 
  plot_layout(ncol = 3, guides = "collect") &  # Collects legends from all subplots into one
  theme(legend.position = "none")  # Places the legend at the bottom

# Display the final plot
print(final_plot)

dev.print(pdf,"~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/SVLEN.lim.pdf",width=9,height=6)

# store file:
write_delim(gfinal_data_list, "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/gfinal_data_list.txt", delim = "\t")

```

```{r Filter sniffles and GraffiTE calls, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)

# data

sfinal_data_list <- read_delim("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sfinal_data_list.txt", delim = "\t", col_names = T)
gfinal_data_list <- read_delim("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/gfinal_data_list.txt", delim = "\t", col_names = T)


head(sfinal_data_list)
head(gfinal_data_list)

# sniffles2
sfinal_data_filtered <- sfinal_data_list %>%
  mutate(avgDP = DP / NS) %>%
  filter(
    F_MISSING <= 0.25,
    AF > 0.05,
    QUAL >= 40,
    case_when(
      # Species == "Aca" ~ avgDP >= 16 & avgDP <= 89,
      Species == "Ass" ~ avgDP >= 19 & avgDP <= 94,
      Species == "Hga" ~ avgDP >= 15 & avgDP <= 82,
      Species == "Hls" ~ avgDP >= 31 & avgDP <= 136,
      Species == "Hrs" ~ avgDP >= 29 & avgDP <= 108,
      Species == "Hti" ~ avgDP >= 10 & avgDP <= 96,
      # Species == "Nps" ~ avgDP >= 17 & avgDP <= 101,
      Species == "Ppr" ~ avgDP >= 15 & avgDP <= 77,
      Species == "Sms" ~ avgDP >= 13 & avgDP <= 85,
      TRUE ~ FALSE  # Exclude unmatched species
    )
  )


# GraffiTE
df_clean <- gfinal_data_list %>%
  filter(!apply(., 1, function(row) any(grepl(",", row))))

gfinal_data_filtered <- df_clean %>%
  mutate(avgDP = DP / NS) %>%
  filter(
    F_MISSING <= 0.25,
    AF > 0.05,
    QUAL >= 178.1014,
    QUAL <= 5000,
    case_when(
      # Species == "Aca" ~ avgDP >= 18 & avgDP <= 198,
      Species == "Ass" ~ avgDP >= 27 & avgDP <= 217,
      Species == "Hga" ~ avgDP >= 22 & avgDP <= 193,
      Species == "Hls" ~ avgDP >= 53 & avgDP <= 277,
      Species == "Hrs" ~ avgDP >= 39 & avgDP <= 217,
      Species == "Hti" ~ avgDP >= 15 & avgDP <= 206,
      # Species == "Nps" ~ avgDP >= 22 & avgDP <= 261,
      Species == "Ppr" ~ avgDP >= 22 & avgDP <= 176,
      Species == "Sms" ~ avgDP >= 17 & avgDP <= 200,
      TRUE ~ FALSE  # Exclude unmatched species
    )
  )

```

#### *Make Figure summarizing everything SVs*

```{r Plot SVs related stats, echo=FALSE, message=FALSE, warning=FALSE}

# Add indicator columns before joining
sfinal_data_filtered <- sfinal_data_filtered %>% mutate(source_sfinal = TRUE)
gfinal_data_filtered <- gfinal_data_filtered %>% mutate(source_gfinal = TRUE)

# Full join by ID and Species
combined <- full_join(sfinal_data_filtered, gfinal_data_filtered,
                      by = c("ID", "Species"), suffix = c(".sfinal", ".gfinal"))

# Annotate based on presence in each dataset
combined <- combined %>%
  mutate(TE_status = case_when(
    !is.na(source_sfinal) & !is.na(source_gfinal) ~ "both",
    !is.na(source_gfinal) & is.na(source_sfinal)  ~ "TE_only",
    is.na(source_gfinal) & !is.na(source_sfinal)  ~ "nonTE_only",
    TRUE                                           ~ NA_character_
  ))

combined <- combined %>%
  mutate(SVTYPE = coalesce(SVTYPE.sfinal, SVTYPE.gfinal)) %>%
  filter(SVTYPE != ".")


combined <- combined  %>% filter(Species != "Nps" &
                                   Species != "Aca")

####################
# Part b: #
####################
# STEP 1: Clean up TE categories and SV types
combined <- combined %>% filter(combined$TE_status != "TE_only") %>%
  mutate(TE_group = case_when(
    TE_status == "nonTE_only" ~ "nonTE",
    TE_status %in% c("TE_only", "both") ~ "TE",
    TRUE ~ NA_character_
  ))

# STEP 2: Count per Species, TE_group, and SVTYPE
sv_counts <- combined %>%
  filter(!is.na(SVTYPE.sfinal) | !is.na(SVTYPE.gfinal)) %>%
  mutate(SVTYPE = coalesce(SVTYPE.sfinal, SVTYPE.gfinal)) %>%
  filter(SVTYPE != ".") %>%  # remove SVTYPE == "."
  group_by(Species, SVTYPE.sfinal) %>% # TE_group
  dplyr::summarise(Count = n(), .groups = "drop")

# STEP 3: Plot with bars per species
sv_counts$Species <- factor(sv_counts$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))

sv_counts <- mutate(sv_counts, reproMode = ifelse(sv_counts$Species == "Ppr" |
                                                # sv_counts$Species == "Nps" |
                                                # sv_counts$Species == "Ona" |
                                                sv_counts$Species == "Hrs" |
                                                sv_counts$Species == "Hls" |
                                                sv_counts$Species == "Ass", "asexual", "sexual"))


a <- ggplot(sv_counts, aes(x = Species, y = Count/1000, fill = SVTYPE.sfinal, colour = NULL)) + # TE_group
  geom_bar(stat = "identity", position = "stack", size = .4) + theme_bw()+
  scale_fill_manual(values = rev(c("#BACC81", "#478C5C", "#5CD85A", "#013A20")))+
  ylab("Number of SVs (x 1K)")+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_classic2() +
  theme(# panel.border = element_blank(),
        # panel.grid.major = element_blank(),
        # panel.grid.minor = element_blank(),
        # axis.line = element_line(colour = "black"),
        # axis.line.x = element_blank(),
        # axis.text.x = element_text(size = 10), # angle = 45, hjust = 1, 
        axis.title.x = element_blank(),
        # axis.ticks.x = element_blank(),
        # legend.title = element_text(size = 10),
        # legend.text = element_text(size = 10),
        # panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        legend.position = "top")+
  ggtitle("")


a
ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/SVcounts_Barplot.pdf", a, width = 3, height = 4)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/SVcounts_Barplot.pdf", a, width = 3, height = 3.5)


# Define species and corresponding genome sizes (in Mb)
species <- c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls")
genome_sizes <- c(184.7, 261.0, 138.7, 352.2, 216.0, 272.0, 100.6)

# Create a data frame
df <- data.frame(Species = factor(species, levels = species), GenomeSize = genome_sizes)

# Merge genome size into your TE dataset
df_out3 <- merge(sv_counts, df, by = "Species", all.x = TRUE)

# Calculate counts normalized per Mb genome
df_norm <- df_out3 %>%
  group_by(Species, SVTYPE.sfinal, reproMode, GenomeSize) %>%
  # summarise(Count = n(), .groups = "drop") %>%
  mutate(Count_perMb = Count / GenomeSize)

# Plot normalized counts
p4 <- ggplot(df_norm, aes(x = Species, y = Count_perMb, fill = SVTYPE.sfinal, color = NULL)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = rev(c("#BACC81", "#478C5C", "#5CD85A", "#013A20")))+
  labs(x = NULL, y = "SV per Mb") +
  # theme_bw() +
  scale_color_manual(
    name = "reproMode",
    values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")
  ) +
  theme_classic2() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/countsPerMb.pdf", p4, width = 2.5, height = 1.8)


# plot by TE non TE:

# STEP 2: Count per Species, TE_group, and SVTYPE
sv_counts <- combined %>%
  filter(!is.na(SVTYPE.sfinal) | !is.na(SVTYPE.gfinal)) %>%
  mutate(SVTYPE = coalesce(SVTYPE.sfinal, SVTYPE.gfinal)) %>%
  filter(SVTYPE != ".") %>%  # remove SVTYPE == "."
  group_by(Species, TE_group, SVTYPE) %>%
  summarise(Count = n(), .groups = "drop")

# STEP 3: Plot with bars per species
sv_counts$Species <- factor(sv_counts$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))

sv_counts <- mutate(sv_counts, reproMode = ifelse(sv_counts$Species == "Ppr" |
                                                # sv_counts$Species == "Nps" |
                                                # sv_counts$Species == "Ona" |
                                                sv_counts$Species == "Hrs" |
                                                sv_counts$Species == "Hls" |
                                                sv_counts$Species == "Ass", "asexual", "sexual"))


b <- ggplot(sv_counts, aes(x = Species, y = Count/1000, fill = SVTYPE, colour = reproMode)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ TE_group, ncol = 1) +
  labs(
    x = "Species",
    y = "Number of SVs (x 1K)"
  ) +
  theme_classic() +
  scale_fill_manual(values = rev(c("#BACC81", "#478C5C", "#5CD85A", "#013A20")))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.line.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.title.x = element_blank(),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "top")+
  ggtitle("")

b

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/SVTYPE_Barplot.pdf", b, width = 3, height = 4)



# plot by TE non TE:

# STEP 2: Count per Species, TE_group, and SVTYPE
sv_counts <- combined %>%
  filter(!is.na(SVTYPE.sfinal) | !is.na(SVTYPE.gfinal)) %>%
  mutate(SVTYPE = coalesce(SVTYPE.sfinal, SVTYPE.gfinal)) %>%
  filter(SVTYPE != ".") %>%  # remove SVTYPE == "."
  group_by(Species) %>%
  summarise(Count = n(), .groups = "drop")

# STEP 3: Plot with bars per species
sv_counts$Species <- factor(sv_counts$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))

sv_counts <- mutate(sv_counts, reproMode = ifelse(sv_counts$Species == "Ppr" |
                                                # sv_counts$Species == "Nps" |
                                                # sv_counts$Species == "Ona" |
                                                sv_counts$Species == "Hrs" |
                                                sv_counts$Species == "Hls" |
                                                sv_counts$Species == "Ass", "asexual", "sexual"))


# Define species and corresponding genome sizes
species <- c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls")
genome_sizes <- c(184.7, 261.0, 138.7, 352.2, 216.0, 272.0, 100.6)

# Create a data frame
df <- data.frame(Species = factor(species, levels = species), GenomeSize = genome_sizes)

sv_counts <- merge(sv_counts, df,
                  by = c("Species"), all.x = TRUE)


c <- sv_counts  %>%
  ggplot(.,aes(x=Count/1000,y=GenomeSize))+
  geom_point(alpha=1, size=2, aes(fill = reproMode, colour = reproMode))+
  stat_cor(method = "s", cor.coef.name = "rho", cex=3, )+
  geom_smooth(method = "glm", se=T,size=1,col="blue",fill="grey80")+
  ylab(expression(paste("Genome size (Mb)")))+
  xlab(expression("Number of SVs (x 1K)"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_classic()+
  theme(# panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        # axis.text.x = element_text(angle = 0, hjust = 1, size = 12),
        axis.text.y = element_text(size = 10),
        legend.position = "none",
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10)) #+ ggtitle("A")


c

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/SVcounts_corr.pdf", c, width = 3, height = 3)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/SVcounts_corr.pdf", c, width = 2.5, height = 1.8)


####################
# Part c: #
####################

df <- mutate(combined, reproMode = ifelse(combined$Species == "Ppr" |
                                                # combined$Species == "Nps" |
                                                # combined$Species == "Ona" |
                                                combined$Species == "Hrs" |
                                                combined$Species == "Hls" |
                                                combined$Species == "Ass", "asexual", "sexual"))

df$Species <- factor(df$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))

# make named vector
p2 <- ggplot(df, aes(x = Species, y = log10(abs(SVLEN.sfinal)), fill = as.factor(reproMode))) +
  geom_violin(aes(fill = reproMode, color = reproMode), alpha=.5, linewidth=.3,  drop = FALSE, scale = "width") +
  geom_boxplot(aes(fill = reproMode), color = "black", width=0.1,alpha=.8, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  theme_classic2() +
  labs(title = NULL,
       x = NULL,
       y = expression(paste("log"[10], " SV length (kb)"))) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
    #panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    axis.text.y = element_text(size = 10),
    legend.position = "none") #+ coord_flip()


p2

ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/SVlengths.pdf", p2, width = 2.5, height = 1.8)


############################
# plot allele frequencies: #
############################

# # Derive MAF column and TE_group label
# combined <- combined %>%
#   mutate(TE_group = case_when(
#     TE_status == "nonTE_only" ~ "nonTE",
#     TE_status %in% c("TE_only", "both") ~ "TE",
#     TRUE ~ NA_character_
#   ))

combined$Species <- factor(combined$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))

combined <- mutate(combined, reproMode = ifelse(combined$Species == "Ppr" |
                                                # combined$Species == "Nps" |
                                                # combined$Species == "Ona" |
                                                combined$Species == "Hrs" |
                                                combined$Species == "Hls" |
                                                combined$Species == "Ass", "asexual", "sexual"))


maf_box <- combined %>% filter(TE_group == "nonTE") %>%
  ggplot(., aes(x = Species, y = `MAF.sfinal`, fill = TE_group)) + theme_bw()+
  # geom_violin(aes(color= reproMode, fill = reproMode), alpha=.4, linewidth=.3, scale = "width") +
  geom_boxplot(aes(color= reproMode, fill = reproMode), color = "black", alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA) +
  labs(
    y = "Minor allele frequency",
    x = NULL,
    fill = "TE Category"
  ) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.line.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.title.x = element_blank(),
        # axis.ticks.x = element_blank(),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "top"
  )

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/nonTE_SV_MAF_Boxplot.pdf", maf_box, width = 3, height = 3)


# Perform Kruskal-Wallis test to compare distributions
sv_data <- combined %>% filter(TE_group == "nonTE") 

kruskal_test <- kruskal.test(`MAF.sfinal` ~ Species, data = sv_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(sv_data$`MAF.sfinal`, sv_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load this
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters


# TEonly
sv_data <- combined %>% filter(TE_group == "TE") 

sv_data$MAF.gfinal <- as.numeric(sv_data$MAF.gfinal)

tmaf_box <- ggplot(sv_data, aes(x = Species, y = (`MAF.gfinal`))) + theme_bw()+
  geom_boxplot(aes(color= reproMode, fill = reproMode), color = "black", alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA) +
  labs(
    y = "Minor allele frequency",
    x = NULL,
    fill = "TE Category"
  ) +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.line.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.title.x = element_blank(),
        # axis.ticks.x = element_blank(),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "top"
  )

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/TE_MAF_Boxplot.pdf", tmaf_box, width = 3, height = 3)


# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test((`MAF.gfinal`) ~ Species, data = sv_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test((sv_data$`MAF.gfinal`), sv_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load this
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters

# Plot heterozygosity:
# Loop through each species and write unique IDs to separate files
unique_species <- unique(combined$Species)

for (sp in unique_species) {
  ids <- unique(combined$ID[combined$Species == sp & combined$TE_group == "TE"])
  writeLines(ids, paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/graffite/unique_ids_", sp, ".txt"))
}


for (sp in unique_species) {
  ids <- unique(combined$ID[combined$Species == sp & combined$TE_group == "nonTE"])
  writeLines(ids, paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sniffles2/unique_ids_", sp, ".txt"))
}

# Het is calculated on motoko

# sfs <- ggplot(combined, aes(x = AF.sfinal, fill = TE_group, color = reproMode)) +
#   geom_histogram(bins = 10, color = "black") +
#   facet_wrap(~Species, scales = "free_y") +
#   labs(x = "Allele frequency", y = "Count") +
#   scale_color_manual(name="reproMode",values = c(
#     "sexual"="#F56B39",
#     "asexual"="#11A7BB"))+
#   theme_bw() +
#   theme_classic()
# 
# sfs
```

```{r SVs het, echo=FALSE, message=FALSE, warning=FALSE}
#############################
# plot SV heterozygosity:
#############################
# Set working directory (change this to your folder)
path = "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/het/"

# SVs:

# List all .het.txt files in the directory
files<-dir(path,pattern ="\\.het.txt") #.nonTEs.het.txt

# Initialize an empty dataframe to store results
heterozygosity_data <- data.frame()

# Loop through each file and read data
for (file in files) {
  # Extract species name from the filename (assuming format: Species.het.txt)
  species_name <- gsub(".het.txt", "", file)
  
  # Read data
  df <- read.table(paste(path, file, sep = ""), header = TRUE, sep = "\t")
  
  # Add species column
  df$Species <- species_name
  
  # Combine data
  heterozygosity_data <- rbind(heterozygosity_data, df)
}


# Simplify Species names (remove .TEs / .nonTEs)
heterozygosity_data <- heterozygosity_data %>%
  mutate(Species = str_replace(Species, "\\.nonTEs|\\.TEs", ""))

# Average per sample
heterozygosity_summary <- heterozygosity_data %>%
  group_by(Sample, Species) %>%
  summarise(
    Heterozygous = sum(Heterozygous, na.rm = TRUE),
    Homozygous = sum(Homozygous, na.rm = TRUE),
    Heterozygosity_Rate = Heterozygous/(Heterozygous + Homozygous),
    .groups = "drop"
  )


# Convert species to a factor
heterozygosity_summary$Species <- as.factor(heterozygosity_summary$Species)

# add reproductive mode
heterozygosity_summary <- mutate(heterozygosity_summary, reproMode = ifelse(heterozygosity_summary$Species == "Ppr" |
                                                                        # heterozygosity_summary$Species == "Nps" |
                                                                        # heterozygosity_summary$Species == "Ona" |
                                                                        heterozygosity_summary$Species == "Hrs" |
                                                                        heterozygosity_summary$Species == "Hls" |
                                                                        heterozygosity_summary$Species == "Ass", "asexual", "sexual"))

# order species:
heterozygosity_summary$Species <- factor(heterozygosity_summary$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  

# Plot boxplot of heterozygosity rate
p <- ggplot(heterozygosity_summary, aes(x = Species, y = Heterozygosity_Rate, fill = Species)) +
  # geom_boxplot(aes(fill = reproMode), color = "black", alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  geom_violin(aes(fill = reproMode, color = reproMode), alpha=.5, linewidth=.3,  drop = FALSE, scale = "width") +
  geom_boxplot(aes(fill = reproMode), color = "black", width=0.1, alpha=.8, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  theme_bw() +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  labs(title = NULL,
       x = NULL,
       y = "SV heterozygosity") +
  theme_classic2() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
    #panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    legend.position = "top")
  # theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
  #       panel.grid.major = element_blank(),
  #       panel.grid.minor = element_blank(),
  #       axis.line = element_line(colour = "black"),
  #       axis.line.x = element_blank(),
  #       axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
  #       axis.title.x = element_blank(),
  #       axis.ticks.x = element_blank(),
  #       legend.title = element_text(size = 10),
  #       legend.text = element_text(size = 10),
  #       legend.position = "top"
  # )

p

# Save plot
ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/het/SVs_Heterozygosity_Boxplot.pdf", p, width = 3, height = 3)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/SVs_Heterozygosity_Boxplot.pdf", p, width = 3, height = 3.5)

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(Heterozygosity_Rate ~ Species, data = heterozygosity_summary)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(heterozygosity_summary$Heterozygosity_Rate, heterozygosity_summary$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load thsi
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters


# TEs:

# List all .het.txt files in the directory
files<-dir(path,pattern ="\\.TEs.het.txt")

# Initialize an empty dataframe to store results
heterozygosity_data <- data.frame()

# Loop through each file and read data
for (file in files) {
  # Extract species name from the filename (assuming format: Species.het.txt)
  species_name <- gsub(".TEs.het.txt", "", file)
  
  # Read data
  df <- read.table(paste(path, file, sep = ""), header = TRUE, sep = "\t")
  
  # Add species column
  df$Species <- species_name
  
  # Combine data
  heterozygosity_data <- rbind(heterozygosity_data, df)
}

# Convert species to a factor
heterozygosity_data$Species <- as.factor(heterozygosity_data$Species)

# add reproductive mode
heterozygosity_data <- mutate(heterozygosity_data, reproMode = ifelse(heterozygosity_data$Species == "Ppr" |
                                                                        # heterozygosity_data$Species == "Nps" |
                                                                        # heterozygosity_data$Species == "Ona" |
                                                                        heterozygosity_data$Species == "Hrs" |
                                                                        heterozygosity_data$Species == "Hls" |
                                                                        heterozygosity_data$Species == "Ass", "asexual", "sexual"))

# order species:
heterozygosity_data$Species <- factor(heterozygosity_data$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  

# Plot boxplot of heterozygosity rate
p <- ggplot(heterozygosity_data, aes(x = Species, y = Heterozygosity_Rate, fill = Species)) +
  geom_boxplot(aes(fill = reproMode), color = "black", alpha=.5, linewidth=.3, position = position_dodge(width = 0.9),outlier.shape = NA)+
  theme_bw() +
  scale_fill_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  labs(title = NULL,
       x = NULL,
       y = "TE heterozygosity") +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.line.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "top"
  )

p

# Save plot
ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/het/TEs_Heterozygosity_Boxplot.pdf", p, width = 3, height = 3)

# Perform Kruskal-Wallis test to compare distributions
kruskal_test <- kruskal.test(Heterozygosity_Rate ~ Species, data = heterozygosity_data)

# Print test result
print(kruskal_test)

# Post-hoc pairwise Wilcoxon test
pwc <- pairwise.wilcox.test(heterozygosity_data$Heterozygosity_Rate, heterozygosity_data$Species, p.adjust.method = "BH")

pwc

# this function should help convert sig into letters:
tri.to.squ<-function(x)
{
  rn<-row.names(x)
  cn<-colnames(x)
  an<-unique(c(cn,rn))
  myval<-x[!is.na(x)]
  mymat<-matrix(1,nrow=length(an),ncol=length(an),dimnames=list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x)==rn[int],colnames(x)==cn[ext]])) next
      mymat[row.names(mymat)==rn[int],colnames(mymat)==cn[ext]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
      mymat[row.names(mymat)==cn[ext],colnames(mymat)==rn[int]]<-x[row.names(x)==rn[int],colnames(x)==cn[ext]]
    }
    
  }
  return(mymat)
}

mymat<-tri.to.squ(pwc$p.value)

# load thsi
library(multcompView)

myletters<-multcompLetters(mymat,compare="<=",threshold=0.05,Letters=letters)

myletters
```

```{r Heterozygosity, echo=FALSE, message=FALSE, warning=FALSE}

library(dplyr)
library(ggplot2)
library(ggpubr)

# Set working directory (change this to your folder)
path = "~/sciebo/Postdoc/sexAsex/popGen/diversity/het/"

# List all .het.txt files in the directory
files<-dir(path,pattern =".txt")

# Initialize an empty dataframe to store results
heterozygosity_data <- data.frame()

# Loop through each file and read data
for (file in files) {
  # Extract species name from the filename (assuming format: Species.het.txt)
  species_name <- gsub(".het.txt", "", file)
  
  # Read data
  df <- read.table(paste(path, file, sep = ""), header = TRUE, sep = "\t")
  
  # Add species column
  df$Species <- species_name
  
  # Combine data
  heterozygosity_data <- rbind(heterozygosity_data, df)
}

# Convert species to a factor
heterozygosity_data$Species <- as.factor(heterozygosity_data$Species)

heterozygosity_data <- heterozygosity_data %>% filter(Species != "Ona"|
                                                        Species != "Osa"|
                                                        Species != "Nps"|
                                                        Species != "Aca")


# add reproductive mode
heterozygosity_data <- mutate(heterozygosity_data, reproMode = ifelse(heterozygosity_data$Species == "Ppr" |
                                                                        heterozygosity_data$Species == "Hrs" |
                                                                        heterozygosity_data$Species == "Hls" |
                                                                        heterozygosity_data$Species == "Ass", "asexual", "sexual"))

# order species:
heterozygosity_data$Species <- factor(heterozygosity_data$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  


# Define folder with .bed files
bed_folder <- "/Users/mohammederrbii/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/haploblocks"

# List .bed files
bed_files <- list.files(path = bed_folder, pattern = "atleast100kb.bed$", full.names = TRUE)

# Read and combine all BED files
haplotype_data <- do.call(rbind, lapply(bed_files, function(file) {
  df <- fread(file, header = FALSE)
  species <- sub("\\..*$", "", basename(file))  # Extract species from filename
  df <- df[, .(Species = species, length = V4)]
  return(df)
}))

# Add reproductive mode using your logic
haplotype_data <- haplotype_data %>%
  mutate(reproMode = ifelse(Species %in% c("Ppr", "Hrs", "Hls", "Ass"), 
                            "asexual", "sexual"))

# Set factor levels for species (for plot order)
haplotype_data$Species <- factor(haplotype_data$Species, levels = c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))


# Step 1: Mean heterozygosity per species
het_means <- heterozygosity_data %>%
  group_by(Species, reproMode) %>%
  summarise(mean_het = mean(Heterozygosity_Rate), .groups = "drop")

# Step 2: Mean haplotype length per species
haplo_means <- haplotype_data %>%
  group_by(Species, reproMode) %>%
  summarise(mean_haplo = mean(length), .groups = "drop")

# Step 3: Merge both summaries
combined_means <- inner_join(het_means, haplo_means, by = c("Species", "reproMode"))

# Step 4: Correlation plot
c <- combined_means %>%
  ggplot(aes(x = mean_het, y = mean_haplo / 1e6)) +
  geom_point(alpha = 1, size = 3, aes(color = reproMode, fill = reproMode), shape = 21) +
  stat_cor(method = "spearman", cor.coef.name = "rho", size = 3) +
  geom_smooth(method = "glm", se = TRUE, color = "blue", fill = "grey80", size = 1) +
  labs(
    x = "Mean heterozygosity",
    y = "Mean haplotype length (Mb)"
  ) +
  scale_color_manual(values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")) +
  scale_fill_manual(values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.position = "none"
  )

# Show plot
print(c)

ggsave("~/sciebo/Postdoc/sexAsex/popGen/phasedVCFs/haploblocks/Haploblocks_Heterozygosity_corr.pdf", c, width = 3, height = 3)


```

#### *LD decay analysis*

```{r LD decay analyses, echo=FALSE, message=FALSE, warning=FALSE}
# Load libraries
library(ggplot2)
library(dplyr)
library(stringr)
library(readr)

# Directory containing LD decay files
indir <- "/Users/mohammederrbii/sciebo/Postdoc/sexAsex/popGen/LDdecay/unphased/PopLDdecay/forR"
pattern <- "Multi.LDdecay.unphased.maf0.2."

# List species files (exclude non-species files like .bin, .pdf, .png)
files <- list.files(indir, pattern = paste0(pattern, "[A-Za-z]+$"), full.names = TRUE)

# Extract species names from filenames
species_names <- str_replace(basename(files), pattern, "")

# Read and combine all species data into one dataframe
ld_data <- bind_rows(lapply(seq_along(files), function(i) {
  df <- read_table(files[i], col_names = T, show_col_types = FALSE)
  df <- df %>% transmute(
    DistanceKb = `#Dist` / 1000,
    R2 = `Mean_r^2`,
    Species = species_names[i]
  )
  return(df)
}))

# Custom colors (adjust to your liking)
species_colors <- c(
  # "Aca" = "red",
  "Ass" = "black",
  "Hga" = "blue",
  "Hls" = "purple",
  "Hrs" = "green",
  "Hti" = "orange",
  # "Nps" = "yellow",
  "Ppr" = "brown",
  "Sms" = "cyan"   # optional if you want Ona
)

# Define asexual species
asexuals <- c("Ppr", "Hrs", "Hls", "Ass")

# Add reproductive mode column
ld_data <- ld_data %>%
  mutate(reproMode = ifelse(Species %in% asexuals, "Asexual", "Sexual"))


# Plot with ggplot + loess
p <- ggplot(ld_data, aes(x = DistanceKb, y = R2, color = Species)) +
  geom_point(alpha = 0.01) +  # raw points
  geom_smooth(method = "loess", span = 0.05, se = T, linewidth = 0.5, alpha = 0.2) + # smoothed curves
  scale_color_manual(values = species_colors) +
  scale_color_manual(values = c("Sexual" = "#F56B39", "Asexual" = "#11A7BB")) +
  labs(
    x = "Distance (Kb)",
    y = expression(r^2),
    color = "Species"
  ) +
  coord_cartesian(xlim = c(0, 300), ylim = c(0, 1)) +
  theme_classic2() +
  theme(legend.position = "right")


# Save plot
# ggsave("~/sciebo/Postdoc/sexAsex/popGen/LDdecay/unphased/PopLDdecay/Multi.LDdecay.maf0.2.ggplot.pdf", p, width = 8, height = 6)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/Multi.LDdecay.maf0.2.ggplot.pdf", p, width = 5, height = 2.5)

# Also preview in RStudio
print(p)

 
# Define asexual species
asexuals <- c("Ppr", "Hrs", "Hls", "Ass")

# Add reproductive mode column
ld_data <- ld_data %>%
  mutate(reproMode = ifelse(Species %in% asexuals, "Asexual", "Sexual"))


ggplot(ld_data, aes(x = DistanceKb, y = R2, color = reproMode)) +
  geom_point(alpha = 0.2, size = .5) +  # raw points
  geom_smooth(method = "loess", span = 0.05, se = FALSE, linewidth = 0.5) +
  scale_color_manual(values = c("Sexual" = "#F56B39", "Asexual" = "#11A7BB")) +
  facet_wrap(~Species) +
  labs(x="Distance (Kb)", y=expression(r^2), color="Repro. mode") +
  theme_classic()


# Estimate half-decay distance

half_decay <- ld_data %>%
  group_by(Species) %>%
  arrange(DistanceKb) %>%
  mutate(baseline   = mean(R2[DistanceKb < 1], na.rm = TRUE),
         half_value = baseline * (2/4)) %>%
  summarise(baseR2 = unique(baseline),
            halfR2 = unique(half_value), 
            {
              idx <- which(R2 <= first(half_value))[1]
              
              if (is.na(idx)) {
                # never reached: assign max distance
                hd <- max(DistanceKb, na.rm = TRUE)
              } else if (idx == 1) {
                # already below half at start
                hd <- min(DistanceKb, na.rm = TRUE)
              } else {
                # linear interpolation
                x1 <- DistanceKb[idx-1]; x2 <- DistanceKb[idx]
                y1 <- R2[idx-1]; y2 <- R2[idx]
                hv <- first(half_value)
                hd <- x1 + (hv - y1) * (x2 - x1) / (y2 - y1)
              }
              
              tibble(half_decay_dist = hd)
            })



# Compare statistically (sexual vs asexual)
half_decay <- half_decay %>%
  mutate(reproMode = ifelse(Species %in% asexuals, "Asexual", "Sexual"))


half_decay
# write_delim(half_decay, "~/sciebo/Postdoc/sexAsex/popGen/LDdecay/unphased/PopLDdecay/half_decay.maf0.2.txt", delim = "\t")
write_delim(half_decay, "~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/half_decay.maf0.2.txt", delim = "\t")


# Boxplot + stats
ggplot(half_decay, aes(x = reproMode, y = (half_decay_dist), fill = reproMode)) +
  geom_violin(alpha=0.6) +
  geom_jitter(width=0.1, size=2) +
  labs(y="Half-decay distance (Kb)", x="Reproductive mode") +
  scale_fill_manual(values = c("Sexual" = "#F56B39", "Asexual" = "#11A7BB")) +
  theme_classic()

# Wilcoxon test
wilcox.test(half_decay_dist ~ reproMode, data=half_decay, p.adjust.method = "bonferoni")

# PGLS analysis:

# Read phylogeny
tree <- read.tree(text="(Hrs:0.25251170135,((Ass:0.0370616018,Sms:0.0279731614):0.2859294992[100],((Hga:0.2068827003,(Nps:0.1293561260,Ppr:0.0921946231):0.0269159387[100]):0.0287608487[100],((Ona:0.0997975298,Osa:0.1187814774):0.1204310951[100],(Aca:0.0851089497,(Hls:0.0308508642,Hti:0.0332963161):0.0797602904[100]):0.1021325842[100]):0.1385490472[100]):0.0932480595[100]):0.25251170135);")

# Ensure names match (drop missing)
tree <- drop.tip(tree, setdiff(tree$tip.label, half_decay$Species))

# p <- ggtree(tree,ladderize=T,right=T,size=1.5,lineend='square') + geom_tiplab(size=3,color=1,offset=.02,fontface = "bold")

half_decay <- half_decay %>% filter(Species %in% tree$tip.label)

# Build comparative data object
half_decay <- as.data.frame(half_decay)
rownames(half_decay) <- half_decay$Species

comp <- comparative.data(
  phy = tree,
  data = half_decay,
  names.col = Species,
  vcv = TRUE,
  na.omit = TRUE,
  warn.dropped = TRUE
)

# Check comp
# summary(comp)


# Run PGLS models
pgls_ld <- pgls(half_decay_dist ~ reproMode, data = comp, lambda = "ML")
summary(pgls_ld)

```

#### *SV analysis: functional annotation*

```{r TE polymorphism, echo=FALSE, message=FALSE, warning=FALSE}
# prepare the data:
gfinal_data_list <- read_delim("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/gfinal_data_list.txt", delim = "\t", col_names = T)

gfinal_data_list <- gfinal_data_list %>% filter(Species != "Aca" &
                                                  Species != "Nps")
# load new freq 
table_files <- list.files(path = "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE/", pattern = "\\.freq\\.frq$", full.names = TRUE)
species_list <- sub("\\.freq\\.frq$", "", basename(table_files))

# Initialize list to store data
data_list <- list()

# Loop over species
for (species in species_list) {
  cat("Processing:", species, "\n")
  
  table_file <- file.path("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE/", paste0(species, ".freq.frq"))
  
  if (!file.exists(table_file)) {
    warning("Missing table file: ", table_file)
    next
  }
  
  # Read the data and assign to list using species name as key
  data_list[[species]] <- read.table(table_file, header = F, sep = "\t", stringsAsFactors = FALSE, skip = 1)
  
  data_list[[species]]$Species <- species

}

# Combine all tables into a single data frame by row binding
freq_data_list <- bind_rows(data_list)
colnames(freq_data_list) <- c("CHROM", "POS", "N_ALLELES", "N_CHR", "Ref_Allele", "Alt_Allele", "Species")


# load annotations 
table_files <- list.files(path = "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE", pattern = "\\.Classes.table$", full.names = TRUE)
species_list <- sub("\\.Classes.table$", "", basename(table_files))

# Initialize list to store data
data_list <- list()

# Loop over species
for (species in species_list) {
  cat("Processing:", species, "\n")
  
  table_file <- file.path("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE/", paste0(species, ".Classes.table"))
  
  if (!file.exists(table_file)) {
    warning("Missing table file: ", table_file)
    next
  }
  
  # Read the data and assign to list using species name as key
  data_list[[species]] <- read.table(table_file, header = F, sep = "\t", stringsAsFactors = FALSE)
  
  data_list[[species]]$Species <- species

}

# Combine all tables into a single data frame by row binding
class_data_list <- bind_rows(data_list)
colnames(class_data_list) <- c("CHROM", "POS", "Class", "Species")

# merge all three:
head(gfinal_data_list)

head(freq_data_list)

head(class_data_list)

combined <- merge(freq_data_list, class_data_list, by = c("CHROM", "POS", "Species")) %>%
  merge(., gfinal_data_list[gfinal_data_list$SVTYPE != ".",], by = c("CHROM", "POS", "Species"))


final_combined <- combined[,c(1,2,6,7,9,11,12,18,3,8)]

final_combined$Species <- factor(final_combined$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  

final_combined <- mutate(final_combined, reproMode = ifelse(final_combined$Species == "Ppr" |
                                                # final_combined$Species == "Nps" |
                                                # final_combined$Species == "Ona" |
                                                final_combined$Species == "Hrs" |
                                                final_combined$Species == "Hls" |
                                                final_combined$Species == "Ass", "asexual", "sexual"))



# fix umbigious annotations:
df_out <- final_combined %>%
  rowwise() %>%
  mutate(
    # split the Class column by commas, trim whitespace
    Class_list = list(str_split(Class, ",")[[1]] %>% str_trim()),
    
    # count how many classes are matched
    n_classes = length(Class_list),
    
    # tabulate classes, pick the most frequent one
    top_class = {
      tab <- sort(table(Class_list), decreasing = TRUE)
      names(tab)[1]
    }
  ) %>%
  ungroup()


df_out1 <- df_out %>% filter(df_out$n_classes == 1) #& df_out$Class != "Unknown"


### 1. Pie charts of top_class composition per species ----------------

cols<-read.table("/Users/mohammederrbii/sciebo/Postdoc/sexAsex/popGen/src/TEcolors.land", sep = "\t")
# Turn your color table into a named vector
col_vector <- setNames(cols$V2, cols$V1)

df_pie <- df_out1 %>%
  group_by(Species, top_class) %>%
  summarise(n = n(), .groups="drop") %>%
  group_by(Species) %>%
  mutate(frac = n / sum(n))


df_new <- df_pie %>%
  mutate(
    # Extract the TE superfamily (everything before the first "/")
    class = str_extract(top_class, "^[^/]+"),
    
    # Keep detailed top_class if it's ≥ 1%, otherwise collapse under _other
    # but keep "Unknown" as-is
    class2 = case_when(
      class == "Unknown" ~ "Unknown",  # leave Unknown unchanged
      frac >= 0.01 & top_class != class ~ top_class,  # keep detailed names
      TRUE ~ paste0(class, "_other")  # group the rest
    )
  )

df_new



df_collapsed <- df_new %>%
  group_by(Species, class2) %>%
  summarise(frac = sum(frac), .groups = "drop")


# Make sure your fill variable matches the names in col_vector (e.g., class2)
p1 <- ggplot(df_collapsed, aes(x = "", y = frac, fill = class2)) +
  geom_col(width = 1, color = "white", size = .2) +
  coord_polar(theta = "y") +
  facet_wrap(~Species, nrow = 7) +
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_void() +
  theme(legend.position = "right",
        legend.key.size = unit(0.5, 'cm')) +
  scale_fill_manual(values = col_vector) +
  guides(fill=guide_legend(ncol=1))

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE/pieChart_collapsed_fractions.pdf", p1, width = 4, height = 6)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/pieChart_collapsed_fractions.pdf", p1, width = 6, height = 10)


### 2. Alt_allele distributions for each species ----------------------
df_out2 <- df_out1 #%>% filter(df_out$Class == "LTR/Gypsy")


df_out2 <- df_out2 %>%
  mutate(
    # Extract the TE superfamily (before first "/")
    Class = str_extract(top_class, "^[^/]+")
  )

# make named vector
col_vector <- setNames(cols$V2, cols$V1)

p2 <- ggplot(df_out2, aes(x = Alt_Allele, fill = Class, color = reproMode)) +
  geom_histogram(bins = 10, color = "black") +
  facet_wrap(~Species, scales = "free_y", nrow=7) +
  labs(x = "Allele frequency", y = "Count") +
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_bw() +
  scale_fill_manual(values = col_vector) +
  theme_classic()



ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE/SFS.pdf", p2, width = 8, height = 6)


# p2 <- ggplot(df_out2, aes(x = Alt_Allele, fill = Class, color = reproMode)) +
#   geom_density(bins = 20, alpha = 0.3, position = "stack") +
#   facet_wrap(~Species, scales = "free_y", nrow=7) +
#   # labs(x = "Allele frequency", y = "Count") +
#   scale_color_manual(name="reproMode",values = c(
#     "sexual"="#F56B39",
#     "asexual"="#11A7BB"))+
#   theme_bw() +
#   scale_fill_manual(values = col_vector) +
#   theme_classic()

p2 <- ggplot(df_out2, aes(x = Alt_Allele, fill = Class, color = reproMode)) +
  geom_histogram(aes(y = ..count.. / 1000), bins = 10, color = NA) +  # divide count by 1000
  facet_wrap(~Species, scales = "free_y", nrow = 7) +
  labs(x = "Allele frequency", y = "Count (x1000)") +  # update label
  scale_color_manual(name = "reproMode", values = c(
    "sexual" = "#F56B39",
    "asexual" = "#11A7BB"
  )) +
  theme_bw() +
  theme(legend.key.size = unit(0.5, 'cm'))+
  scale_fill_manual(values = col_vector) +
  theme_classic()

ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/TE.SFS.pdf", p2, width = 3, height = 7)


p3 <- ggplot(df_out2, aes(x = Species, fill = Class, color = reproMode)) +
  geom_bar() +
  labs(x = "Species", y = "Count") +
  theme_bw() +
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_fill_manual(values = col_vector) +
  theme_classic()

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE/counts.pdf", p3, width = 4, height = 4)


# Define species and corresponding genome sizes (in Mb)
species <- c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls")
genome_sizes <- c(184.7, 261.0, 138.7, 352.2, 216.0, 272.0, 100.6)

# Create a data frame
df <- data.frame(Species = factor(species, levels = species), GenomeSize = genome_sizes)

# Merge genome size into your TE dataset
df_out3 <- merge(df_out2, df, by = "Species", all.x = TRUE)

# Calculate counts normalized per Mb genome
df_norm <- df_out3 %>%
  group_by(Species, Class, reproMode, GenomeSize) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Count_perMb = Count / GenomeSize)

# Plot normalized counts
p4 <- ggplot(df_norm, aes(x = Species, y = Count_perMb, fill = Class, color = reproMode)) +
  geom_col(position = "stack") +
  labs(x = "Species", y = "TE counts per Mb genome") +
  theme_bw() +
  scale_color_manual(
    name = "reproMode",
    values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")
  ) +
  scale_fill_manual(values = col_vector) +
  theme_classic()

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/TE/countsPerMb.pdf", p4, width = 6, height = 4)

### 3. Bin allele frequencies into one figure -------------------------
# Bin allele frequencies
df_binned <- df_out %>%
  mutate(freq_bin = case_when(
    Alt_Allele < 0.1 ~ "Low-frequency (<0.1)",
    Alt_Allele >= 0.1 & Alt_Allele <= 0.9 ~ "Common (0.1–0.9)",
    Alt_Allele > 0.9 ~ "Fixed (>0.9)"
  ))

# Count per species and frequency bin
df_summary <- df_binned %>%
  group_by(Species, freq_bin) %>%
  summarise(n = n(), .groups="drop")

# Plot stacked bar chart
p5 <- ggplot(df_summary, aes(x = Species, y = n, fill = freq_bin)) +
  geom_bar(stat="identity", position="fill") +   # use 'fill' for proportions
  labs(title = "TE frequency categories by species",
       y = "Proportion of TEs", x = "Species", fill = "Frequency bin") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p5

```

```{r non TE structural variants, echo=FALSE, message=FALSE, warning=FALSE}
# Required libraries
library(dplyr)
library(stringr)
library(ggvenn)   # nice venn plots
library(purrr)

# List of species
species <- c("Ass","Hga","Hls","Hrs","Hti","Ppr","Sms")

# Directory containing your files
sv_dir <- "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/"

# Function to read a given species' SV annotation files
read_sv_annotations <- function(sp) {
  
  # file paths
  files <- list(
    exonic = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.exons")),
    genic  = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.genes")),
    intron = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.introns")),
    inter  = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.intergenic"))
  )
  
  # read safely (handle intergenic with only 1 column)
  read_file <- function(path, category) {
    if (file.exists(path)) {
      df <- read.table(path, sep="\t", fill=TRUE, stringsAsFactors=FALSE)
      df$category <- category
      colnames(df)[1:2] <- c("SVid","Gene")
      return(df)
    } else {
      return(data.frame(SVid=character(), Gene=character(), category=character()))
    }
  }
  
  # bind everything
  dat <- bind_rows(
    read_file(files$exonic, "exonic"),
    read_file(files$genic,  "genic"),
    read_file(files$intron, "intronic"),
    read_file(files$inter,  "intergenic")
  )
  
  dat$species <- sp
  return(dat)
}

# ---- Load all species data
sv_all <- map_df(species, read_sv_annotations)

sv_all <- sv_all[,-3]


# ---- 1. Get counts of unique SVs per category per species
summary_counts <- sv_all %>%
  group_by(species, category) %>%
  summarise(n_unique = n_distinct(SVid), .groups="drop")

# ---- 2. Count intron-only SVs (present in intronic but not exonic for each species)
introns_only <- sv_all %>%
  group_by(species) %>%
  summarise(
    n_unique = n_distinct(SVid[category == "intronic" & !(SVid %in% SVid[category == "exonic"])]),
    .groups = "drop"
  ) %>%
  mutate(category = "intronic")

# ---- 3. Count exon-only SVs (present in exonic but not intronic for each species)
exons_only <- sv_all %>%
  group_by(species) %>%
  summarise(
    n_unique = n_distinct(SVid[category == "exonic" & !(SVid %in% SVid[category == "intronic"])]),
    .groups = "drop"
  ) %>%
  mutate(category = "exonic")


# ---- 4. Replace intronic counts in summary_counts
summary_counts_replaced <- summary_counts %>%
  filter(category != "intronic") %>%     # keep everything except intronic
  bind_rows(introns_only)                # add the corrected intronic counts

# ---- 5. Replace exonic counts in summary_counts
summary_counts_replaced <- summary_counts_replaced %>%
  filter(category != "exonic") %>%     # keep everything except exonic
  bind_rows(exons_only)                # add the corrected exonic counts

# (optional) arrange for nice output
summary_counts_replaced <- summary_counts_replaced %>%
  arrange(species, category)

print(summary_counts_replaced)

# ---- 2. Make Venn diagrams for each species
# Example: plot for Ppr
make_venn <- function(sp) {
  subset <- sv_all %>% filter(species == sp)
  
  venn_list <- list(
    # intergenic= unique(subset$SVid[subset$category=="intergenic"]),
    # genic     = unique(subset$SVid[subset$category=="genic"]),
    exonic    = unique(subset$SVid[subset$category=="exonic"]),
    intronic  = unique(subset$SVid[subset$category=="intronic"])
  )
  
  # plot venn
  p <- ggvenn(venn_list, fill_color = c("#3399ff","#66cc99","#ff6666","#999999"),
              stroke_size = 0.5, set_name_size = 4) + ggtitle(sp)
  
  # ggsave(paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/", sp, "_SV_categories.venn.pdf"), plot=p, width=3, height=2)
  ggsave(paste0("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/", sp, "_SV_categories.venn2.pdf"), plot=p, width=3, height=2)
  return(p)
}

# run venns for all species
venn_plots <- map(species, make_venn)

# 
```

```{r no TE SVs, echo=FALSE, message=FALSE, warning=FALSE}
# prepare the data:
sfinal_data_list <- read_delim("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/outputs/sfinal_data_list.txt", delim = "\t", col_names = T)

sfinal_data_list <- sfinal_data_list %>% filter(Species != "Aca" &
                                                  Species != "Nps")
# load new freq 
table_files <- list.files(path = "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/", pattern = "\\.freq\\.frq$", full.names = TRUE)
species_list <- sub("\\.freq\\.frq$", "", basename(table_files))

# Initialize list to store data
data_list <- list()

# Loop over species
for (species in species_list) {
  cat("Processing:", species, "\n")
  
  table_file <- file.path("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/", paste0(species, ".freq.frq"))
  
  if (!file.exists(table_file)) {
    warning("Missing table file: ", table_file)
    next
  }
  
  # Read the data and assign to list using species name as key
  data_list[[species]] <- read.table(table_file, header = F, sep = "\t", stringsAsFactors = FALSE, skip = 1)
  
  data_list[[species]]$Species <- species

}

# Combine all tables into a single data frame by row binding
freq_data_list <- bind_rows(data_list)
colnames(freq_data_list) <- c("CHROM", "POS", "N_ALLELES", "N_CHR", "Ref_Allele", "Alt_Allele", "Species")

freq_data_list <- unique(freq_data_list)

# combined <- merge(freq_data_list, sfinal_data_list[sfinal_data_list$SVTYPE != ".",], by = c("CHROM", "POS", "Species"))


# load annotations 
# List of species
species <- c("Ass","Hga","Hls","Hrs","Hti","Ppr","Sms")

# Directory containing your files
sv_dir <- "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/"

# Function to read a given species' SV annotation files
read_sv_annotations <- function(sp) {
  
  # file paths
  files <- list(
    exonic = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.exons")),
    genic  = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.genes")),
    intron = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.introns")),
    inter  = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.intergenic"))
  )
  
  # read safely (handle intergenic with only 1 column)
  read_file <- function(path, category) {
    if (file.exists(path)) {
      df <- read.table(path, sep="\t", fill=TRUE, stringsAsFactors=FALSE)
      df$category <- category
      colnames(df)[1:2] <- c("ID","Gene")
      return(df)
    } else {
      return(data.frame(SVid=character(), Gene=character(), category=character()))
    }
  }
  
  # bind everything
  dat <- bind_rows(
    read_file(files$exonic, "exonic"),
    read_file(files$genic,  "genic"),
    read_file(files$intron, "intronic"),
    read_file(files$inter,  "intergenic")
  )
  
  dat$Species <- sp
  return(dat)
}

# ---- Load all species data
sv_all <- map_df(species, read_sv_annotations)

sv_all <- sv_all[,-3]

# merge all three:
head(sfinal_data_list)

head(freq_data_list)

head(sv_all)


combined <- merge(sfinal_data_list[sfinal_data_list$SVTYPE != ".",], sv_all, by = c("ID", "Species")) %>%
  merge(., freq_data_list, by = c("CHROM", "POS", "Species"))



# set levels and reproductive mode:
combined$Species <- factor(combined$Species, levels = rev(rev(c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls"))))  

final_combined <- mutate(combined, reproMode = ifelse(combined$Species == "Ppr" |
                                                # combined$Species == "Nps" |
                                                # combined$Species == "Ona" |
                                                combined$Species == "Hrs" |
                                                combined$Species == "Hls" |
                                                combined$Species == "Ass", "asexual", "sexual"))



### Pie charts of top_class composition per species ----------------

# ---- 1. Get counts of unique SVs per category per species
summary_counts <- final_combined %>%
  group_by(Species, category) %>%
  summarise(n_unique = n_distinct(ID), .groups="drop")

# ---- 2. Count intron-only SVs (present in intronic but not exonic for each species)
introns_only <- final_combined %>%
  group_by(Species) %>%
  summarise(
    n_unique = n_distinct(ID[category == "intronic" & !(ID %in% ID[category == "exonic"])]),
    .groups = "drop"
  ) %>%
  mutate(category = "intronic")

# ---- 3. Count exon-only SVs (present in exonic but not intronic for each species)
exons_only <- final_combined %>%
  group_by(Species) %>%
  summarise(
    n_unique = n_distinct(ID[category == "exonic" & !(ID %in% ID[category == "intronic"])]),
    .groups = "drop"
  ) %>%
  mutate(category = "exonic")

# ---- 4. Count exon-only SVs (present in exonic but not intronic for each species)
both <- final_combined %>%
  group_by(Species) %>%
  summarise(
    n_unique = n_distinct(ID[category == "exonic" & (ID %in% ID[category == "intronic"])]),
    .groups = "drop"
  ) %>%
  mutate(category = "exointro")


# ---- 5. Replace intronic counts in summary_counts
summary_counts_replaced <- summary_counts %>%
  filter(category != "intronic") %>%     # keep everything except intronic
  bind_rows(introns_only)                # add the corrected intronic counts

# ---- 6. Replace exonic counts in summary_counts
summary_counts_replaced <- summary_counts_replaced %>%
  filter(category != "exonic") %>%     # keep everything except exonic
  bind_rows(exons_only)                # add the corrected exonic counts

# ---- 7. Replace exonic counts in summary_counts
summary_counts_replaced <- summary_counts_replaced %>%
  bind_rows(both)                # add the  intronic/exonic counts


# arrange for nice output
summary_counts_replaced <- summary_counts_replaced %>%
  arrange(Species, category)


df_pie <- summary_counts_replaced %>% filter(category != "genic") %>%
  group_by(Species, category) %>%
  group_by(Species) %>%
  mutate(frac = n_unique / sum(n_unique))



# Make sure your fill variable matches the names in col_vector (e.g., class2)
p1 <- ggplot(df_pie, aes(x = "", y = frac, fill = category)) +
  geom_col(width = 1, color = "white", size = .2) +
  coord_polar(theta = "y") +
  facet_wrap(~Species) +
  # add text labels outside
  geom_text(
    aes(label = round(frac*100,2), color = category),
    color = "white", size = 3, fontface = "bold"
  ) +
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_void() +
  theme(legend.position = "right") +
  scale_fill_manual(values = c("#3399ff", "#ff6666", "#999999", "#66cc99", "orange"))



ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/pieChart_fractions.pdf", p1, width = 10, height = 6)
ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/tmp/pieChart_fractions.pdf", p1, width = 10, height = 6)



### Alt_allele distributions for each species ----------------------

# make named vector
p2 <- ggplot(final_combined %>% filter(category == "intergenic" |
                                         category == "genic"), aes(x = Alt_Allele, fill = category, color = reproMode)) +
  geom_histogram(aes(y = ..count.. / 1000), bins = 10, color = NA) +  # divide count by 1000
  facet_wrap(~Species, nrow = 7, scales = "free_y") +
  labs(x = "Allele frequency", y = "Count (x1000)") +  # update label
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  theme_bw() +
  scale_fill_manual(values = c("orange", "darkblue", "#3399ff", "#ff6666", "#999999", "#66cc99")) +
  scale_y_continuous(
    labels = function(x) sprintf("%.0f", x)  # ❗ removes decimals
  ) +
  theme_classic()

# ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SFS.pdf", p2, width = 4, height = 14)

# p2 <- ggplot(final_combined %>% 
#                filter(category %in% c("intergenic", "genic")), 
#              aes(x = Alt_Allele)) +
#   geom_histogram(aes(y = ..count../1000, fill = category, color = NA), bins = 10, color = NA, alpha = 0.6) +  # histogram
#   geom_density(aes(weight = ..count../1000, color = reproMode), size = 1, alpha = 0.8) +    # trend line
#   facet_wrap(~Species, nrow = 7, scales = "free_y") +
#   labs(x = "Allele frequency", y = "Count (x1000)") +
#   scale_color_manual(name="reproMode", values = c(
#     "sexual"="#F56B39",
#     "asexual"="#11A7BB")) +
#   scale_fill_manual(values = c("orange", "#999999")) +
#   theme_classic()

ggsave("~/sciebo/Postdoc/sexAsex/popGen/manuscript/figures/nonTE.SFS.pdf", p2, width = 3.05, height = 7)



df_out2 <- merge(final_combined %>% filter(category == "intergenic" |
                                         category == "genic"), df, by = "Species", all.x = TRUE)


p3 <- ggplot(df_out2, aes(x = Species, fill = category, color = reproMode)) +
  geom_bar() +
  labs(x = "Species", y = "Count") +
  theme_bw() +
  scale_color_manual(name="reproMode",values = c(
    "sexual"="#F56B39",
    "asexual"="#11A7BB"))+
  scale_fill_manual(values = c("orange", "#999999", "#3399ff", "#ff6666", "#999999", "#66cc99"))+
  theme_classic()

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/counts.pdf", p3, width = 4, height = 4)


# Define species and corresponding genome sizes (in Mb)
species <- c("Hrs", "Sms", "Ass", "Hga", "Ppr", "Hti", "Hls")
genome_sizes <- c(184.7, 261.0, 138.7, 352.2, 216.0, 272.0, 100.6)

# Create a data frame
df <- data.frame(Species = factor(species, levels = species), GenomeSize = genome_sizes)

# Merge genome size into your TE dataset
df_out3 <- merge(final_combined %>% filter(category == "intergenic" |
                                         category == "genic"), df, by = "Species", all.x = TRUE)

# Calculate counts normalized per Mb genome
df_norm <- df_out3 %>%
  group_by(Species, category, reproMode, GenomeSize) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Count_perMb = Count / GenomeSize)

# Plot normalized counts
p4 <- ggplot(df_norm, aes(x = Species, y = Count_perMb, fill = category, color = reproMode)) +
  geom_col(position = "stack") +
  labs(x = "Species", y = "TE counts per Mb genome") +
  theme_bw() +
  scale_color_manual(
    name = "reproMode",
    values = c("sexual" = "#F56B39", "asexual" = "#11A7BB")
  ) +
  scale_fill_manual(values = c("orange", "#999999", "#3399ff", "#ff6666", "#999999", "#66cc99"))+
  theme_classic()

ggsave("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/countsPerMb.pdf", p4, width = 6, height = 4)

```

```{r non TE SVs: topGO, echo=FALSE, message=FALSE, warning=FALSE}
# Required libraries

# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# 
# BiocManager::install("topGO")

# Load libraries
library(topGO)
library(plyr)
library(tidyverse)
library(dplyr)
library(stringr)
library(purrr)

# List of species
species <- c("Ass","Hga","Hls","Hrs","Hti","Ppr","Sms")

# Directory containing your files
sv_dir <- "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/"

# Function to read a given species' SV annotation files
read_sv_annotations <- function(sp) {
  
  # file paths
  files <- list(
    exonic = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.exons")),
    genic  = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.genes")),
    intron = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.introns")),
    inter  = file.path(sv_dir, paste0(sp, ".sniffles2.CHR_POS_END_ID.table.sorted.intergenic"))
  )
  
  # read safely (handle intergenic with only 1 column)
  read_file <- function(path, category) {
    if (file.exists(path)) {
      df <- read.table(path, sep="\t", fill=TRUE, stringsAsFactors=FALSE)
      df$category <- category
      colnames(df)[1:2] <- c("SVid","Gene")
      return(df)
    } else {
      return(data.frame(SVid=character(), Gene=character(), category=character()))
    }
  }
  
  # bind everything
  dat <- bind_rows(
    read_file(files$exonic, "exonic"),
    read_file(files$genic,  "genic"),
    read_file(files$intron, "intronic"),
    read_file(files$inter,  "intergenic")
  )
  
  dat$species <- sp
  return(dat)
}

# ---- Load all species data
sv_all <- map_df(species, read_sv_annotations)

sv_all <- sv_all[,-3]

# extract genes affected only by exonic SVs
genes_exonic_list <- lapply(species, function(sp) {
  unique(sv_all$Gene[sv_all$category == "exonic" & sv_all$species == sp])
})

# Name the list elements by species for clarity
names(genes_exonic_list) <- species

# length(genes_exonic_list$Sms)


##########################################################################
# GO enrichment analysis of genes under intense/relaxed selection in Hrs #
##########################################################################


# Path setup
go_dir <- "~/sciebo/Postdoc/sexAsex/genomes/geneAnno/GOterms"
out_dir <- "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO"

# Loop over all species
for (sp in species) {
  message("=== Processing species: ", sp, " ===")
  
  # Step 1: Read and reformat GO universe ---------------------
  # infile <- file.path(go_dir, paste0(sp, ".universe.GO.forR.txt"))
  outfile <- file.path(go_dir, paste0(sp, ".universe.GO.forR.final.txt"))
  
  # if (!file.exists(infile)) {
  #   warning("GO file not found for ", sp, ", skipping...")
  #   next
  # }
  # 
  # a <- read.csv(infile, sep = "\t", header = FALSE)
  # colnames(a) <- c("V1", "V2")
  # 
  # all <- ddply(a, .(V1), summarize,
  #              V2 = paste(unique(V2), collapse = ","))
  # 
  # write_delim(all, outfile, col_names = FALSE, delim = "\t")
  
  # Step 2: Load GO mappings -----------------------------------
  geneID2GO <- readMappings(file = outfile)
  geneUniverse <- names(geneID2GO)
  
  # Step 3: Define genes of interest ----------------------------
  if (!sp %in% names(genes_exonic_list)) {
    warning("No gene list for ", sp, ", skipping...")
    next
  }
  
  genesOfInterest <- genes_exonic_list[[sp]]
  
  geneList <- factor(as.integer(geneUniverse %in% genesOfInterest))
  names(geneList) <- geneUniverse
  
  # Step 4: Create topGOdata objects ----------------------------
  GOdata_MF <- new("topGOdata", ontology = "MF", allGenes = geneList,
                   annot = annFUN.gene2GO, gene2GO = geneID2GO)
  GOdata_BP <- new("topGOdata", ontology = "BP", allGenes = geneList,
                   annot = annFUN.gene2GO, gene2GO = geneID2GO)
  GOdata_CC <- new("topGOdata", ontology = "CC", allGenes = geneList,
                   annot = annFUN.gene2GO, gene2GO = geneID2GO)
  
  # Step 5: Run tests -------------------------------------------
  resultFis_MF <- runTest(GOdata_MF, algorithm = "weight01", statistic = "fisher")
  resultFis_BP <- runTest(GOdata_BP, algorithm = "weight01", statistic = "fisher")
  resultFis_CC <- runTest(GOdata_CC, algorithm = "weight01", statistic = "fisher")
  
  # Step 6: Extract tables --------------------------------------
  table_MF <- GenTable(GOdata_MF, classicFisher = resultFis_MF,
                       orderBy = "resultFisher", ranksOf = "resultFisher", topNodes = 200)
  table_BP <- GenTable(GOdata_BP, classicFisher = resultFis_BP,
                       orderBy = "resultFisher", ranksOf = "resultFisher", topNodes = 200)
  table_CC <- GenTable(GOdata_CC, classicFisher = resultFis_CC,
                       orderBy = "resultFisher", ranksOf = "resultFisher", topNodes = 200)
  
# Step 7: Adjust p-values and filter --------------------------
  table_MF$FDR <- p.adjust(as.numeric(table_MF$classicFisher), method = "fdr")
  table_BP$FDR <- p.adjust(as.numeric(table_BP$classicFisher), method = "fdr")
  table_CC$FDR <- p.adjust(as.numeric(table_CC$classicFisher), method = "fdr")
  
  MF <- subset(table_MF, as.numeric(classicFisher) < 0.05)
  BP <- subset(table_BP, as.numeric(classicFisher) < 0.05)
  CC <- subset(table_CC, as.numeric(classicFisher) < 0.05)
  
  MF$ontology <- "MF"
  BP$ontology <- "BP"
  CC$ontology <- "CC"
  
  # Step 8: Combine and write results ---------------------------
  res <- rbind(MF, BP, CC)
  res$Species <- sp
  res$classicFisher <- as.numeric(res$classicFisher)
  
  outfile_res <- file.path(out_dir, paste0(sp, ".GO.enrichment.exonicSVs.txt"))
  write.table(res, outfile_res, quote = FALSE, sep = "\t", row.names = FALSE, dec = ".")
}


```

```{r non TE SVs topGO summary, echo=FALSE, message=FALSE, warning=FALSE}

# Define your output directory
out_dir <- "~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO"

# List all the output files (e.g., enrichment result tables)
# You can filter for only .txt files, if that’s what you have
files <- list.files(path = out_dir, pattern = "\\.txt$", full.names = TRUE)

# Read all files into a list of data frames
all_go_results <- lapply(files, function(f) {
  df <- read.delim(f, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  # df$Species <- sub("\\.GO.enrichment.exonicSVs.txt$", "", basename(f))  # Add species name based on filename
  df
})

# Combine them all into one large data frame
combined_go <- do.call(rbind, all_go_results)

# Check the structure
str(combined_go)

# Preview
head(combined_go)



# Label species by reproductive mode
combined_go$ReproMode <- ifelse(combined_go$Species %in% 
                                  c("Ppr", "Nps", "Ona", "Hrs", "Hls", "Ass"), 
                                "Asexual", "Sexual")
# get sig GO terms
sig_go <- combined_go %>%
  filter(classicFisher < 0.01) %>%
  distinct(GO.ID, Term, ontology, classicFisher, Species, ReproMode)


# Visualize GOs:

# if (!requireNamespace("BiocManager", quietly=TRUE))
#   install.packages("BiocManager")
# BiocManager::install("simplifyEnrichment", force = TRUE)
# library(simplifyEnrichment)


# simplifyGO(unique(go_id), ont="MF")



# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install(c("rrvgo", "GOSemSim", "org.Dm.eg.db"))  # use org.Dm.eg.db or your organism


library(rrvgo)
library(GOSemSim)
library(org.Dm.eg.db) # or org.Dm.eg.db
library(dplyr)


species_list <- unique(combined_go$Species)

# species_list <- c("Ppr", "Hga")

for (sp in species_list) {
  cat("Processing:", sp, "\n")
  
  # Filter to significant GO terms (e.g., FDR < 0.05)
  df <- combined_go %>% 
    filter(Species == sp & ontology == "BP", classicFisher < 0.01)
  
  if (nrow(df) == 0) next
  
  # Compute semantic similarity matrix
  simMatrix <- calculateSimMatrix(df$GO.ID,
                                  orgdb = "org.Dm.eg.db",
                                  ont = "BP",  # Change to BP, MF or CC as needed
                                  method = "Rel")
  
  # Reduce redundancy (keep representative terms)
  # Create a named score vector
  scores <- setNames(-log10(df$classicFisher), df$GO.ID)
  
  # Now reduce redundancy
  reducedTerms <- reduceSimMatrix(simMatrix,
                                  scores = scores,
                                  threshold = 0.7,
                                  orgdb = "org.Dm.eg.db")
  
  # Save reduced list
  write.csv(reducedTerms, 
            paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO/BP_GO_reduced_classicFisher_", sp, ".csv"), 
            row.names = FALSE)
  
  # Plot treemap
  pdf(paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO/BP_GO_treemap_classicFisher_", sp, ".pdf"))
  treemapPlot(reducedTerms)
  dev.off()
  
  # Plot scatter
  p <- scatterPlot(simMatrix, reducedTerms,addLabel=T,onlyParents=F,size="size")
  ggsave(filename = paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO/BP_GO_scatter_classicFisher_", sp, ".pdf"),
         plot = p,
         width = 5, height = 5)
}


# compare the two confirmed asex to the sexual ones:
# get GO id:
go_id = sig_go[sig_go$Species == "Ppr" & sig_go$ontology == "MF"|
                       sig_go$Species == "Hls" & sig_go$ontology == "MF",]


# deduplicate GO terms (GO.ID) 
go_id_filtered <- go_id %>%
  group_by(GO.ID) %>%                           # group by GO term
  dplyr::arrange(classicFisher, .by_group = TRUE) %>%  # sort each group by FDR ascending
  dplyr::slice(1) %>%                                  # keep only the lowest-FDR entry per GO.ID
  ungroup()                                     # remove grouping



# Compute semantic similarity matrix
simMatrix <- calculateSimMatrix(go_id_filtered$GO.ID,
                                orgdb = "org.Dm.eg.db",
                                ont = "MF",  # Change to BP, MF or CC as needed
                                method = "Rel")


# Create a named score vector
scores <- setNames(-log10(go_id_filtered$classicFisher), go_id_filtered$GO.ID)

# Now reduce redundancy
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores = scores,
                                threshold = 0.7,
                                orgdb = "org.Dm.eg.db")

# # Save reduced list
# write.csv(reducedTerms, 
#           paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO/MF_GO_reduced_", sp, ".csv"), 
#           row.names = FALSE)

# Plot treemap
pdf(paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO/MF_GO_treemap_classicFisher_Ppr_Hls.pdf"))
treemapPlot(reducedTerms)
dev.off()

# Plot scatter
p <- scatterPlot(simMatrix, reducedTerms,addLabel=T,onlyParents=F,size="size")
ggsave(filename = paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO/MF_GO_scatter_classicFisher_Ppr_Hls_v1.pdf"),
       plot = p,
       width = 5, height = 5)

df.n10<-scatterPlot(simMatrix,reducedTerms,addLabel=T,onlyParents=F,size="size")$data

# labelSize <- 1.5
# p.n10<-ggplot2::ggplot(df.n10, ggplot2::aes(x = V1, y = V2, color = parentTerm)) +
#   ggplot2::geom_point(ggplot2::aes(size = size), alpha = 0.5) +
#   ggplot2::scale_color_discrete(guide = "none") + ggplot2::scale_size_continuous(range = c(0, 20)) + ggplot2::scale_x_continuous(name = "") +
#   ggplot2::scale_y_continuous(name = "") + ggplot2::theme_minimal() +
#   ggplot2::theme(
#     axis.text.x = ggplot2::element_blank(),
#     axis.text.y = ggplot2::element_blank()
#   ) + geom_text_repel(
#   data          = subset(df.n10, V1 > 0 & parent == rownames(df.n10)),
#   aes(label         = parentTerm),
#   nudge_x       = 1.4 - subset(df.n10, V1 > 0 & parent == rownames(df.n10))$V1,
#   segment.size  = 0.2,
#   segment.color = "grey50",
#   direction     = "y",
#   hjust         = 0
# ) +
#   geom_text_repel(
#     data          = subset(df.n10, V1 <= 0 & parent == rownames(df.n10)),
#     aes(label         = parentTerm),
#     nudge_x       = -1.2 - subset(df.n10, V1 <= 0 & parent == rownames(df.n10))$V1,
#     segment.size  = 0.2,
#     segment.color = "grey50",
#     direction     = "y",
#     hjust         = 1
#   )+coord_cartesian(xlim=c(-10,10))

# p.n11 <- ggplot(df.n10, aes(x = V1, y = V2, color = parentTerm)) +
#   geom_point(aes(size = size), alpha = 0.5) +
#   scale_color_discrete(guide = "none") +
#   scale_size_continuous(range = c(1, 10)) +
#   theme_minimal() +
#   theme(
#     axis.text.x = element_blank(),
#     axis.text.y = element_blank(),
#     panel.grid = element_blank()
#   ) +
#   geom_text_repel(
#     data = df.n10 %>% dplyr::distinct(parentTerm, .keep_all = TRUE),
#     # data = df.n10 %>% dplyr::filter(size > quantile(size, 0.9)),
#     aes(label = parentTerm),
#     size = 3,
#     segment.size = 0.2,
#     segment.color = "grey50",
#     direction = "both",
#     max.overlaps = 1000
#   ) +
#   coord_cartesian(xlim = c(-10, 10)) +
#   labs(x = NULL, y = NULL)


p.n12 <- ggplot(df.n10, aes(x = V1, y = V2, color = parentTerm)) +
  geom_point(aes(size = size), alpha = 0.5) +
  scale_color_discrete(guide = "none") +
  scale_size_continuous(range = c(1, 10)) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(10, 10, 10, 10)
  ) +
  geom_text_repel(
    data = df.n10 %>% dplyr::distinct(parentTerm, .keep_all = TRUE),
    aes(label = str_wrap(parentTerm, width = 25)),  # ✅ Wrap text after 25 chars
    size = 3,
    segment.size = 0.2,
    segment.color = "grey50",
    direction = "both",
    max.overlaps = 1000,
    box.padding = 0.5,
    force = 3
  ) +
  coord_cartesian(xlim = c(-10, 10)) +
  labs(x = NULL, y = NULL)


ggsave(filename = paste0("~/sciebo/Postdoc/sexAsex/popGen/graffiTE/filteredVCFs/nonTE/SVannotation/topGO/MF_GO_scatter_classicFisher_Ppr_Hls_v2.pdf"),
       plot = p.n12,
       width = 5, height = 5)

```
